%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large Installation}

{\tiny (Ref: https://rasa.com/docs/core/installation/)}
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{Prerequisites}
\begin{itemize}
\item Python
\item Spacy
\item Rasa Starter Pack
\item Rasa NLU
\item Rasa Core
\item Slack credentials (for Slack as front-end)
\end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{Python}
\begin{itemize}
\item Install Anaconda 4.2.0 for Python 3.5 or Ananconda 5.2.0 for Python 3.6 
\item Else from Python.org and then additionally numpy and scipy
\item Ubuntu : 
\begin{lstlisting}
sudo apt-get install build-essential python-dev git
\end{lstlisting}
\item Windows Build tools: Make sure the Microsoft $VC++$ Compiler Visual Studio 2015 is installed, so python can compile any dependencies or https://visualstudio.microsoft.com/visual-cpp-build-tools/ Download the installer and select VC++ Build tools in the list.
\end{itemize}

(Note: On main page of Anaconda is https://www.anaconda.com/distribution/ you see Python 3.7, which is problematic, gives HTTP error while creating env. I had to delete 3.7 and get 3.6. Btw, this site does not mention which Anaconda installer has which Python version. WHY? thats the most crucial info. Anyway, I have given them here.)

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{Python Environment}
\begin{itemize}
\item By installing conda, you get base or the root environment, which is the default.
\item Practical tip: DO NOT install any packages in the root. ALWAYS create and env and install inside the new env.
\item Env is needed especially for fragile packages like Python (its treated as a package) and rasa.
\item So, \lstinline|conda create -n myenv python=3.6| 
\end{itemize}

(Note: Env management is again a sour point. It creates complete copy (deeeep) of python 3.6 and all other packages inside "envs" folder. Goes to 1.5 GB!! Can someone optimize it?)
\end{frame}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{Spacy}
If not done already, Spacy Installation: https://spacy.io/usage/
\begin{itemize}
\item Pip
\begin{lstlisting}
pip install -U spacy
\end{lstlisting}
\item Conda
\begin{lstlisting}
conda install -c conda-forge spacy
\end{lstlisting}
\item (Optional) Download English model (in CMD terminal with Administrator permission)
\begin{lstlisting}
python -m spacy download en
\end{lstlisting}
\item Validate
\begin{lstlisting}
python -m spacy validate
\end{lstlisting}
\end{itemize}
\end{frame}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{RASA Starter Pack}
\begin{itemize}
\item Clone the starter pack provided by Rasa from https://github.com/RasaHQ/starter-pack-rasa-stack
\item It gives latest running application
\item Let's take a look at the folder structure and the files that were created 
\item Install git, can git clone above package.
\item \lstinline|cd starter-pack-rasa-stack;activate rasa; pip install -r requirements.txt|

\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{RASA Starter Pack}
\begin{itemize}
\item The ``domain.yml'' file describes the travel assistant's domain. It specifies the list of intents, entities, slots, and response templates that the assistant understands and operates with.
\item The ``data/nlu\_data.md'' file describes each intent with a set of examples that are then fed to Rasa NLU for training.
\item The ``data/stories.md'' file provides Rasa with sample conversations between users and the travel assistant that it can use to train its dialog management model.
\item Rasa provides a lot flexibility in terms of configuring the NLU and core components. For now, we'll use the default ``nlu\_config.yml'' for NLU and ``policies.yml'' for the core model.
\end{itemize}


\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{Pipelines}
\begin{itemize}
\item RASA-NLU is made up of a few components, each doing some specific work (intent detection, entity extraction, etc.). 
\item Each component may have some specific dependencies and installations. 
\item Options like MITIE (NLP + ML), Spacy and Sklearn are available to choose from. 
\item We will be using Spacy-Sklearn here.
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{Rasa NLU and Core}
\begin{itemize}
\item Rasa NLU  with spacy\_sklearn pipeline (Most recent github, https://rasa.com/docs/nlu/installation/)
\begin{lstlisting}
pip install rasa_nlu[spacy]
\end{lstlisting}
\item Rasa Core (https://rasa.com/docs/core/installation/)
\begin{lstlisting}
pip install -U rasa_core==0.9.6
\end{lstlisting}
Note: we are not using the latest. Trying latest as well.
% \item Language Model
% \begin{lstlisting}
% python -m spacy download en_core_web_md
% python -m spacy link en_core_web_md en --force;
% \end{lstlisting}
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{Alternatives}
\begin{itemize}
\item To use the tensorflow\_embedding pipeline you will need to install tensorflow as well as the scikit-learn and sklearn-crfsuite libraries. 
\begin{lstlisting}
pip install rasa_nlu[tensorflow]
\end{lstlisting}
\item MITIE : The MITIE backend performs well for small datasets, but training can take very long if you have more than a couple of hundred examples.
\begin{lstlisting}
pip install git+https://github.com/mit-nlp/MITIE.git
pip install rasa_nlu[mitie]
\end{lstlisting}
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{Rasa Core}
\begin{itemize}
\item Latest (Most recent github)
\begin{lstlisting}
git clone https://github.com/RasaHQ/rasa_nlu.git
cd rasa_nlu
pip install -r requirements.txt
pip install -e .
\end{lstlisting}
\item Ready installation
\begin{lstlisting}
pip install rasa_core
\end{lstlisting}
\end{itemize}
\end{frame}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{UI/Front-End}
\begin{itemize}
\item Client UI can be a web page (using frameworks like Flask in Python) or a mobile app. 
\item Flask is simple to code and runs locally. 
\begin{lstlisting}
pip install flask
\end{lstlisting}
\item More info at https://www.tutorialspoint.com/flask/
\item Many UI front-ends are possible, such as Facebook, Slack, etc. They need API based credentials for fetching what user types and posting the answers generated by Chatbot.
\end{itemize}

\end{frame}

