

\begin{enumerate}
\item \label{stringval} Try to find more solutions to the vibrating string problem $\partial^2 y/\partial t^2=\partial^2 y/\partial x^2$
using the ansatz
$$
y(x,t)=\sin(\omega t) f(x)\, . 
$$
What equation must $f(x)$ obey? Can you write this as an eigenvector equation? Suppose that the string has length $L$ and $f(0)=f(L)=0$.
Can you find any solutions for $f(x)$?

\item Let $M=\begin{pmatrix}
2 & 1 \\
0 & 2
\end{pmatrix}$.  Find all eigenvalues of $M$.  Does $M$ have two \hyperref[indendent]{linearly independent} eigenvectors?  
Is there a basis in which the matrix of $M$ is diagonal? ({\it I.e.}, can $M$ be diagonalized?)

\phantomnewpage

\item Consider $L \colon \Re^2\rightarrow \Re^2$ with $$L\colvec{x\\y}
=\colvec{x\cos \theta +y\sin \theta\\ -x\sin \theta + y\cos \theta}\, .$$
\begin{enumerate}
\item Write the matrix of $L$ in the basis $\colvec{1\\0}, \colvec{0\\1}$.

\item When $\theta\neq 0$, explain how $L$ acts on the plane.  Draw a picture.

\item Do you expect $L$ to have invariant directions? (Consider also special values of $\theta$.)

\item Try to find real eigenvalues for $L$ by solving the equation 
\[
L(v)=\lambda v.
\]

\item Are there complex eigenvalues for $L$, assuming that $i=\sqrt{-1}$ exists?
\end{enumerate}

\phantomnewpage

\item Let $L$ be the linear transformation $L \colon \Re^3\rightarrow \Re^3$ given by 
$$L\colvec{x\\y\\z}=\colvec{ x+y \\  x+z \\ y+z}.$$  
Let $e_i$ be the vector with a one in the $i$th position and zeros in all other positions.
\begin{enumerate}
\item Find $Le_i$ for each $i=1,2,3$.
\item Given a matrix $M=\begin{pmatrix}
m^1_1 & m^1_2 & m^1_3\\
m^2_1 & m^2_2 & m^2_3\\
m^3_1 & m^3_2 & m^3_3\\
\end{pmatrix}$, what can you say about $Me_i$ for each $i$?
\item Find a $3\times 3$ matrix $M$ representing $L$.  
%Choose three nonzero vectors pointing in different directions and show that $Mv=Lv$ for each of your choices.
\item Find the eigenvectors and eigenvalues of \(M.\)

\end{enumerate}

\phantomnewpage

\item Let $A$ be a matrix with eigenvector $v$ with eigenvalue $\lambda$. Show that $v$ is also an eigenvector for $A^2$ and find the corresponding eigenvalue. How about for $A^n$ where $n \in \mathbb{N}$? Suppose that $A$ is invertible. Show that $v$ is also an eigenvector for $A^{-1}$.

\item A \emph{projection}\index{Projection} is a linear operator $P$ such that $P^2 = P$. Let $v$ be an eigenvector with eigenvalue $\lambda$ for a projection $P$, what are all possible values of $\lambda$? Show that every projection $P$ has at least one eigenvector.

Note that every complex matrix has at least 1 eigenvector, but you need to prove the above for \emph{any} field.


\item Explain why the characteristic polynomial of an $n\times n$ matrix has degree~$n$.  Make your explanation easy to read by starting with some simple examples, and then use properties of the determinant to give a \emph{general} explanation.

\phantomnewpage

\item Compute the characteristic polynomial $P_M(\lambda)$ of the matrix $$M=\begin{pmatrix}
a & b \\
c & d
\end{pmatrix}\, .$$  Now, since we can evaluate polynomials on square matrices, we can plug $M$ into its characteristic polynomial and find the {\it matrix} $P_M(M)$.  What do you find from this computation?
Does something similar hold for $3 \times 3$ matrices? (Try assuming that the matrix of $M$ is diagonal to answer this.)
%What about \(n \times n\) matrices ?

\phantomnewpage

\item\label{ddd} {\it Discrete dynamical system.} Let \(M\) be the matrix given by \[M=
\begin{pmatrix}
3 & 2 \\
2 & 3 \\
\end{pmatrix}.\]
Given any vector \(v(0)=
\begin{pmatrix}
x(0) \\
y(0) \\
\end{pmatrix},\) we can create an infinite sequence of vectors \(v(1), v(2), v(3),\) and so on using the rule:
\[v(t+1)=M v(t) \text{ for all natural numbers } t.\] (This is known as a {\it discrete dynamical system} whose {\it initial condition} is \(v(0).\))
\begin{enumerate}
\item Find all eigenvectors and eigenvalues of \(M.\)
\item Find all vectors \(v(0)\) such that \[v(0)=v(1)=v(2)=v(3)=\cdots\] (Such a vector is known as a {\it fixed point} of the dynamical system.)
\item Find all vectors \(v(0)\) such that \(v(0), v(1), v(2), v(3), \ldots\) all point in the same direction. (Any such vector describes an {\it invariant curve} of the dynamical system.)
\end{enumerate}
\Videoscriptlink{eigenvalues_and_eigenvectors_ii_hint_2.mp4}{Hint}{scripts_eigenvalues_and_eigenvectors_ii_hint}



\end{enumerate}

\phantomnewpage
