%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}

\begin{center}
{\Large Claude Code}
\end{center}
\end{frame}

\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Claude Code: Rapid Growth}
  \begin{itemize}
    \item \textbf{4\% of all public GitHub commits} are now authored by Claude Code as of February 2026 --- around 135,000 commits every day.
    \item That represents a \textbf{42,896$\times$ growth} in 13 months since the research preview launched.
    \item \textbf{90\% of Anthropic's own production code} is now AI-written, with Claude Code as the primary tool.
    \item \textbf{8 million+ MCP server downloads} by April 2025 --- up from 100K in November 2024.
    \item Grew from a research preview into the \textbf{de-facto standard} for agentic terminal coding.
    \item Used by engineering teams at Stripe, Notion, and leading AI-native startups for production workflows.
  \end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{What is Claude Code?}
  \begin{itemize}
    \item Agentic coding assistant that lives in your \textbf{terminal}, built by Anthropic.
    \item \textbf{IDE Integration:} Works with VS Code, JetBrains (PyCharm, IntelliJ, WebStorm, etc.), and any terminal-supporting editor.
    \item \textbf{GitHub Integration:} Automate issues, PRs, and code review directly from the CLI.
    \item \textbf{Claude-Powered:} Backed by Anthropic's Claude models (Haiku, Sonnet, Opus) with access to a \textbf{200K token context window} (up to 1M in beta with Opus 4.6+).
      \begin{itemize}
        \item \textit{Enterprise:} AWS Bedrock, Google Vertex AI, Microsoft Foundry
        \item \textit{Direct:} Anthropic Console (API billing) or Claude Pro/Max subscription
      \end{itemize}
    \item \textbf{Safety-First Design:} Fine-grained permission system; enterprise managed settings that cannot be bypassed.
    \item \textbf{Built by Anthropic:} First-party tool with direct access to frontier Claude models and the latest safety research.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Capabilities}
Think of it as having a senior developer sitting next to you
  \begin{itemize}
    \item Writing and debugging code across an entire codebase
    \item Understanding complex project architectures
    \item Refactoring and optimization across multiple files
    \item Answering technical questions with project context
    \item Automating repetitive tasks end-to-end
    \item Managing git workflows: branches, commits, PRs
    \item Connecting to external services (databases, GitHub, Slack) via MCP
  \end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Key Features}
  \begin{itemize}
    \item \textbf{Interactive REPL:} Terminal-native conversation loop; keyboard-driven with intuitive shortcuts.
    \item \textbf{Contextual Understanding:} Reads your entire repo, understands dependencies, structure, languages.
    \item \textbf{Agentic Workflows:} Plans changes, searches files, executes shell commands, modifies code autonomously.
    \item \textbf{Plan / Auto-Accept Modes:} Review proposed changes before applying; \texttt{Shift+Tab} cycles through modes.
    \item \textbf{Slash Commands:} \texttt{/init}, \texttt{/help}, \texttt{/model}, \texttt{/compact}, \texttt{/cost}, and more.
    \item \textbf{File References:} \texttt{@file:path} for scoped queries; paste images directly into the terminal.
    \item \textbf{Auto Memory:} Claude builds and updates \texttt{CLAUDE.md} automatically as it learns your project.
    \item \textbf{Extended Thinking:} Toggle with \texttt{Alt+T} for deep reasoning on complex problems.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Core Integration}
  \begin{itemize}
    \item \textbf{MCP (Model Context Protocol):} Plug in external tools and data sources (GitHub, PostgreSQL, Slack, Jira, and 300+ more) to extend agent capabilities.
    \item \textbf{Hooks System:} 9 lifecycle events (PreToolUse, PostToolUse, Stop, SessionStart, etc.) for deterministic automation. Hooks always execute --- unlike prompts.
    \item \textbf{Subagents:} Spawn specialist agents in isolated contexts (up to 10 in parallel); they return only summaries, preventing context bloat.
    \item \textbf{Skills:} Lazy-loaded \texttt{SKILL.md} folders that encode domain expertise. Loaded only when relevant --- not on every prompt.
    \item \textbf{Plugins:} Installable bundles combining skills, subagents, hooks, and MCP configs into one shareable unit.
    \item \textbf{Permission System:} Fine-grained \texttt{allow}/\texttt{deny}/\texttt{ask} per tool; enterprise managed settings for compliance.
  \end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Claude Code vs Other Tools}
\begin{center}
\begin{tabular}{|p{3cm}|p{1cm}|p{1cm}|p{1.5cm}|}
\hline
\textbf{Feature} & \textbf{Claude Code} & \textbf{OpenCode} & \textbf{Cursor Copilot} \\
\hline
First-party frontier model & \checkmark & -- & -- \\
Hooks (3 handler types) & \checkmark & -- & Limited \\
Plugin system & \checkmark & -- & -- \\
Subagents (up to 10 parallel) & \checkmark & \checkmark & -- \\
MCP support & \checkmark & \checkmark & Limited \\
IDE integration (native) & \checkmark & \checkmark & \checkmark \\
Local / offline models & -- & \checkmark & \checkmark \\
Model-agnostic (75+ models) & -- & \checkmark & -- \\
Open source & -- & \checkmark & -- \\
1M token context & \checkmark & -- & -- \\
Enterprise (Bedrock/Vertex) & \checkmark & -- & Limited \\
\hline
\end{tabular}
\end{center}

{\tiny Claude Code is the \textbf{official} Anthropic tool; deepest integration with Claude models.}
\end{frame}


\section{Summary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Why Use Claude Code?}
  \begin{itemize}
    \item \textbf{Frontier Model Quality:} First-party access to Claude Sonnet and Opus --- the same models powering Anthropic's own 90\%-AI-written codebase.
    \item \textbf{Safety-First Design:} Built-in permission system, enterprise managed settings, and hooks that enforce rules deterministically rather than probabilistically.
    \item \textbf{Hooks Guarantee Execution:} Formatting, security checks, test runs --- hooks always fire at lifecycle events. Prompts rely on the model; hooks do not.
    \item \textbf{Subagents Prevent Context Bloat:} Route exploration and specialized work to isolated agents; the main conversation stays clean and focused.
    \item \textbf{Enterprise-Ready:} AWS Bedrock, Google Vertex AI, Microsoft Foundry, corporate proxy support, managed settings, and data residency options.
    \item \textbf{Terminal-Native Speed:} Faster context-switch than IDE plugins; integrates directly into your existing shell workflow and CI/CD pipelines.
    \item \textbf{Plugin Ecosystem:} Growing community of distributable workflow packages. Build once, share everywhere.
    \item \textbf{Bottom Line:} Faster dev loop, deterministic guardrails, enterprise-grade safety, AI that works \textit{with} your terminal and your team.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% References
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Resources}
  \begin{itemize}
    \item GitHub: \href{https://github.com/anthropics/claude-code}{\texttt{github.com/anthropics/claude-code}}
    \item Docs: \href{https://code.claude.com/docs}{\texttt{code.claude.com/docs}}
    \item Overview: \href{https://docs.claude.com/en/docs/claude-code/overview}{\texttt{docs.claude.com/en/docs/claude-code/overview}}
    \item Awesome Claude Code (community plugins, skills, agents): \href{https://github.com/hesreallyhim/awesome-claude-code}{\texttt{github.com/hesreallyhim/awesome-claude-code}}
    \item Claude Code How-To Guide: \href{https://github.com/luongnv89/claude-howto}{\texttt{github.com/luongnv89/claude-howto}}
    \item Blake Crosley, ``Claude Code CLI: The Complete Guide'', \href{https://blakecrosley.com/guides/claude-code}{\texttt{blakecrosley.com}}, 2026.
    \item Muneeb Ahmad, ``Claude Code Extensions Explained: Skills, MCP, Hooks, Subagents, Agent Teams \& Plugins'', Medium, Mar 2026.
    \item Himanshu Bhoir, ``The 15-File Setup That Turned Claude Code Into My Development Team'', Towards AI, Mar 2026.
    \item MCP Server Registry: \href{https://github.com/modelcontextprotocol/servers}{\texttt{github.com/modelcontextprotocol/servers}}
  \end{itemize}
\end{frame}
