%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}

\begin{center}
{\Large OpenWork by diffeernt-ai}
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{What is OpenWork?}
  \begin{itemize}
    \item \textbf{One-Line Definition:} An open-source, local-first desktop AI agent framework — a guided-workflow GUI and control surface for knowledge workers, built on top of OpenCode.
    \item \textbf{Origin Story:} Created by different-ai as a flexible, model-agnostic alternative to proprietary agentic tools like Claude Code, designed to put users in control of their own AI workflows.
    \item \textbf{What It Does:} Instead of responding one prompt at a time, OpenWork \textit{plans, executes, and delivers} finished work — automating file manipulation, task orchestration, and multi-step agent workflows.
    \item \textbf{Built on OpenCode:} OpenWork acts as a thin, user-friendly GUI layer on top of OpenCode (opencode.ai), with UI actions mapping directly to OpenCode API calls.
    \item \textbf{Status:} Open-source (MIT License) · Desktop app powered by Tauri (Rust + Web) · Available on GitHub at \href{https://github.com/different-ai/openwork}{\texttt{github.com/different-ai/openwork}}.
    \item \textbf{Market Signal:} OpenWork launched in January 2026, weeks after Claude Code, and rapidly drew developer attention as the leading open-source agentic desktop alternative.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{OpenWork vs. Claude Cowork vs. Accomplish AI}
  \begin{itemize}
    \item \textbf{Claude Cowork (Anthropic):} Proprietary · Anthropic models only · Cloud-centric · No source access. Targets non-developers via a polished GUI.
    \item \textbf{OpenWork — different-ai:} Open-source (MIT) · Model-agnostic · Local-first · Direct OpenCode SDK integration. Targets developers and technical users.
    \item \textbf{Accomplish AI (formerly accomplish-ai/openwork):} Open-source · Spawns OpenCode CLI as a subprocess via Electron · Aimed at non-technical users · Rebranded to Accomplish.ai.
    \item \textbf{Key Differentiator:} OpenWork by different-ai integrates directly with the OpenCode SDK (not a subprocess), offering the highest architectural fidelity to OpenCode's design philosophy.
    \item \textbf{Ejectable:} Power users can drop from the GUI to the underlying OpenCode CLI at any time — no lock-in.
    \item \textbf{Tech Stack:} Tauri (Rust + web frontend) — lightweight, native, and cross-platform; versus Electron (Node.js) used by Accomplish AI.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Core Philosophy \& Design Principles}
  \begin{itemize}
    \item \textbf{Local-First, Cloud-Ready:} OpenWork runs on your machine. Workflows execute locally by default; cloud connectivity is opt-in, not mandatory.
    \item \textbf{Model Agnostic:} Bring your own API keys. Switch freely between local models, OpenAI, DeepSeek, or any OpenCode-compatible provider — no ecosystem lock-in.
    \item \textbf{Parity with OpenCode:} Every UI action maps directly to an OpenCode API call. There are no hidden intermediary layers or proprietary abstractions.
    \item \textbf{Transparency:} Audit Logs capture agent activity. File tracking shows exactly what the agent read, wrote, or modified during a task.
    \item \textbf{User Approval for Dangerous Actions:} Plans, permissions, and progress are exposed to the user. Potentially destructive operations require explicit approval.
    \item \textbf{Ejectable Design:} If the GUI is insufficient for your use case, switch to OpenCode CLI — your configuration and workflow definitions transfer seamlessly.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Getting Started — Installation \& Setup}
  \begin{itemize}
    \item \textbf{Prerequisites:} Node.js and pnpm (for building); OpenCode installed from \href{https://opencode.ai}{\texttt{opencode.ai}}.
    \item \textbf{Step 1 — Install OpenCode:} Configure your LLM provider(s) and API keys within OpenCode. OpenWork inherits all provider settings from OpenCode automatically.
    \item \textbf{Step 2 — Launch OpenWork:} Open the desktop app. Select your operating mode:
      \begin{itemize}
        \item \textit{Host Mode:} Runs OpenCode locally and acts as the server for agent execution.
        \item \textit{Client Mode:} Connects to a trusted, running host via URL, token, or QR code — useful for remote or shared setups.
      \end{itemize}
    \item \textbf{Step 3 — Configure Providers:} Set up local models (e.g., Ollama), OpenAI, DeepSeek, or any supported provider. Switch at any time without reinstalling.
    \item \textbf{Development Workflow:} Use \texttt{pnpm dev} to start the full dev stack. Ensure all submodules are updated with \texttt{git submodule update --init --recursive}.
    \item \textbf{Building:} \texttt{pnpm build} compiles the Tauri app for your platform.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{How OpenWork Executes Tasks}
  \begin{itemize}
    \item \textbf{Step 1 — Task Intake:} Type your goal into the chat interface. OpenWork passes it to OpenCode, which analyzes the request and generates a structured execution plan.
    \item \textbf{Step 2 — Plan Review \& Approval:} The plan is surfaced in the UI. You review each proposed step before execution begins — no silent automation.
    \item \textbf{Step 3 — Execution:} OpenCode executes the plan using its tool surface (file read/write, shell commands, browser automation, external integrations).
    \item \textbf{Step 4 — File Handling:} Attach files directly or provide local file paths. The agent can read, edit, summarize, and generate project documents autonomously.
    \item \textbf{Step 5 — Deliver \& Track:} Outputs land in your specified location. The Audit Log records every action taken, every file touched, and every decision made.
    \item \textbf{Commands:} Frequently used plans can be saved as reusable \textit{Commands} — re-run parameterized workflows without re-prompting from scratch.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Key Capabilities}
  \begin{itemize}
    \item \textbf{Local File Access:} Read, write, create, and organize files on your machine — no upload/download cycle, no cloud dependency.
    \item \textbf{Model Flexibility:} Switch between local models (Ollama), OpenAI GPT-4o, DeepSeek Reasoner, or any OpenCode-compatible provider per task.
    \item \textbf{Skills Marketplace:} Extend functionality with pre-built skills — browser automation (Chrome MCP), scheduling, external tool integration, and more.
    \item \textbf{DeepSeek Integration:} For complex tasks requiring multi-step reasoning, OpenWork can delegate to DeepSeek Reasoner — including pivoting from browser automation to local script generation mid-task.
    \item \textbf{Audit Log:} Full, tamper-evident record of agent actions. Know exactly what ran, what changed, and when.
    \item \textbf{Reusable Commands:} Package common workflows as saved commands for instant re-execution with new parameters.
    \item \textbf{OpenWork Apps (Tauri SDK):} A framework within the app for developers to build native applications using OpenCode and Tauri — extending OpenWork as a platform, not just a tool.
    \item \textbf{60fps UI:} Architecture enforces a non-blocking main thread; heavy computations run off-thread to keep the interface responsive.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Audit Log \& Transparency}
  \begin{itemize}
    \item \textbf{What the Audit Log Captures:}
      \begin{itemize}
        \item Every file read, written, or deleted during agent execution
        \item Every shell command or external API call made
        \item Timestamps and model decisions at each step
        \item Plan approvals and user interventions
      \end{itemize}
    \item \textbf{Why It Matters:} Unlike proprietary agentic tools, OpenWork exposes the full execution trace. There are no black-box operations.
    \item \textbf{File Tracking:} The UI surfaces a live file-change feed during task execution. See what the agent is doing in real time.
    \item \textbf{Dangerous Action Gate:} Actions that could cause data loss (e.g., file deletion, irreversible writes) require explicit user \texttt{Allow} confirmation before proceeding.
    \item \textbf{Compliance Advantage:} For teams that cannot send data to proprietary cloud agents, OpenWork's local-first + audit trail design provides a practical path to deploying AI automation within existing data governance policies.
    \item \textbf{Open Source Guarantee:} Because the codebase is MIT-licensed, any organization can audit the agent's behavior at the source level — not just through logs.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Skills Marketplace \& Extensibility}
  \begin{itemize}
    \item \textbf{What Skills Are:} Pre-built capability packages that extend OpenWork's default tool surface. Install a skill to add a new category of agent actions.
    \item \textbf{Available Skill Categories:}
      \begin{itemize}
        \item \textit{Browser Automation} — Chrome MCP integration for web scraping, form filling, and UI-level automation
        \item \textit{Scheduling} — trigger workflows on a time-based schedule
        \item \textit{External Tool Integration} — connect to APIs, webhooks, and third-party services
        \item \textit{Script Generation} — fallback from browser automation to local scripts for robust task completion
      \end{itemize}
    \item \textbf{OpenWork Apps:} The Tauri SDK within OpenWork lets developers build \textit{native applications} on top of OpenCode + Tauri — treating OpenWork as a platform.
    \item \textbf{Plugin Install Flow:}
      \begin{enumerate}
        \item Open OpenWork → navigate to \texttt{Skills} panel
        \item Browse or search the marketplace
        \item Click \texttt{Install} → skill is immediately available in the agent's tool surface
      \end{enumerate}
    \item \textbf{Community-Driven:} Skills are community-contributed and open-source — inspect, fork, or publish your own.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Model Agnosticism — Bring Your Own Keys}
  \begin{itemize}
    \item \textbf{Supported Providers (via OpenCode):}
      \begin{itemize}
        \item \textit{Local:} Ollama (run any open-weight model fully on-device — no API key needed)
        \item \textit{OpenAI:} GPT-4o, GPT-4o-mini, and compatible models
        \item \textit{DeepSeek:} DeepSeek Chat and DeepSeek Reasoner for complex reasoning tasks
        \item \textit{Any OpenCode-compatible provider} — configuration is inherited directly from OpenCode
      \end{itemize}
    \item \textbf{Per-Task Model Selection:} Choose the most appropriate (or cost-effective) model for each task without reconfiguring the app.
    \item \textbf{Cost Control:} Use local models (zero API cost) for routine file operations; switch to frontier models only for tasks that require it.
    \item \textbf{No Vendor Lock-In:} If a provider's pricing changes or a better model is released, switch without migrating your workflows or data.
    \item \textbf{DeepSeek Reasoner Use Case:} For complex tasks requiring multi-hop reasoning or dynamic strategy changes (e.g., ``if browser automation fails, write a local script instead''), DeepSeek Reasoner is the recommended model within OpenWork.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{OpenWork vs. Proprietary Agents — Feature Comparison}
  \begin{center}
  \begin{tabular}{|l|c|c|}
    \hline
    \textbf{Feature} & \textbf{OpenWork (different-ai)} & \textbf{Claude Cowork} \\
    \hline
    Open Source & Yes (MIT) & No \\
    Model Support & Agnostic (Local / API) & Anthropic Only \\
    Primary Use & Desktop Agent / Workflow & Codebase \& Knowledge Work \\
    Execution & Local-First & Cloud-Centric \\
    Audit Log & Yes (full trace) & No \\
    Ejectable to CLI & Yes (OpenCode CLI) & No \\
    Pricing & Free (BYOK) & Paid plans only \\
    Tech Stack & Tauri (Rust / Web) & Proprietary \\
    Skills / Plugins & Community Marketplace & 11 Anthropic Plugins \\
    \hline
  \end{tabular}
  \end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Demo 1 — First Launch \& Setup}
  \begin{itemize}
    \item \textbf{Goal:} Show onboarding flow and provider configuration.
    \item \textbf{Step 1:} Install OpenCode from \href{https://opencode.ai}{\texttt{opencode.ai}}; add your API key(s) to OpenCode config.
    \item \textbf{Step 2:} Clone and build OpenWork (see command below).
    \item \textbf{Step 3:} Launch OpenWork → select \texttt{Host} mode → confirm OpenCode server is running.
    \item \textbf{Step 4:} Navigate to the provider selector — switch from OpenAI to a local Ollama model to demonstrate model agnosticism.
    \item \textbf{What to Show:} The agent tool surface loading, the Audit Log panel initializing, and the Skills Marketplace listing available extensions.
    \item \textbf{Key Message:} Full local setup, zero vendor dependency, and a live audit trail — from clone to running agent in minutes.
  \end{itemize}
\begin{lstlisting}
git clone https://github.com/different-ai/openwork
cd openwork
git submodule update --init --recursive
pnpm install && pnpm build
\end{lstlisting}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Demo 2 — File Automation Task}
  \begin{itemize}
    \item \textbf{Goal:} Show local file access, plan approval, and transparent execution.
    \item \textbf{Setup:} Place a folder of mixed documents (PDFs, CSVs, markdown notes) in a local directory.
    \item \textbf{Task Prompt:}
    \item \textbf{What Happens:} OpenWork generates a structured plan. User reviews and approves. Agent reads files, writes summaries, generates CSV, creates index — Audit Log captures every step.
    \item \textbf{Key Message:} The plan is visible before execution, every file touch is logged, and no data ever leaves the local machine.
  \end{itemize}
\begin{lstlisting}
> Read all files in ~/openwork-demo.
  Summarize each document in one paragraph,
  extract any numerical data into a CSV,
  and produce a master index file with links
  to each summary.
\end{lstlisting}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Demo 3 — Reusable Command Workflow}
  \begin{itemize}
    \item \textbf{Goal:} Show the Commands feature for repeatable, parameterized workflows.
    \item \textbf{Setup:} A recurring task — weekly competitive intelligence report from a list of URLs.
    \item \textbf{First Run — Define the Workflow:}
    \item \textbf{Save as Command:} After the first successful run, save the plan as the command \texttt{/weekly-intel}.
    \item \textbf{Subsequent Runs:} Type \texttt{/weekly-intel} → approve the pre-built plan → done. No re-prompting required.
    \item \textbf{Key Message:} OpenWork learns your recurring workflows. Commands turn one-time tasks into repeatable, auditable automation — without writing any code.
  \end{itemize}
\begin{lstlisting}
> Fetch the pages in urls.txt, extract key
  product announcements and pricing changes
  from the past 7 days, and write a structured
  Markdown report saved to reports/.
\end{lstlisting}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Demo 4 — Browser Automation via Skills}
  \begin{itemize}
    \item \textbf{Goal:} Show the Skills Marketplace extending OpenWork beyond local file operations.
    \item \textbf{Setup:} Install the \textit{Browser Automation (Chrome MCP)} skill from the marketplace.
    \item \textbf{Task Prompt:}
    \item \textbf{What Happens:} OpenWork uses the Chrome MCP skill to launch a controlled browser session, navigates each page, extracts structured data, and writes the CSV — switching to local script generation if a page blocks automation.
    \item \textbf{What to Show:} The Audit Log capturing each browser action, the skill badge appearing in the task trace, and the final CSV delivered locally.
    \item \textbf{Key Message:} Skills turn OpenWork into a general-purpose automation platform without sacrificing auditability or local-first execution.
  \end{itemize}
\begin{lstlisting}
> Go to the pricing pages for the top 5 AI
  coding tools listed in competitors.txt.
  Extract plan names, prices, and feature
  limits. Save results to pricing_table.csv.
\end{lstlisting}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Summary: Why OpenWork Matters}
  \begin{itemize}
    \item \textbf{Paradigm Shift:} From ``AI that answers questions'' to ``AI that executes workflows'' — open, auditable, and local-first.
    \item \textbf{Open Source by Design:} MIT license. No vendor lock-in. Inspect, fork, and deploy on your own terms. Community-driven skills and plugins.
    \item \textbf{Model Freedom:} Run entirely on local models (zero API cost), switch to frontier models when needed — all with the same UI and workflow definitions.
    \item \textbf{Transparent Execution:} Every agent action is logged, every plan is shown before execution, and every dangerous operation requires explicit approval.
    \item \textbf{Built for Extensibility:} The Skills Marketplace and OpenWork Apps SDK let developers build on top of OpenWork as a platform, not just use it as a tool.
    \item \textbf{Closest to OpenCode's Philosophy:} Direct SDK integration (not a subprocess wrapper) means OpenWork reflects OpenCode's architecture faithfully — a GUI that enhances rather than abstracts.
    \item \textbf{The Bigger Picture:} OpenWork is the open-source answer to ``What does AI automation look like when \textit{you} own the stack?''
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Resources \& Links}
  \begin{itemize}
    \item \textbf{GitHub Repository:} \href{https://github.com/different-ai/openwork}{\texttt{github.com/different-ai/openwork}}
    \item \textbf{OpenCode (Agent Engine):} \href{https://opencode.ai}{\texttt{opencode.ai}}
    \item \textbf{License:} MIT — \href{https://github.com/different-ai/openwork/blob/main/LICENSE}{\texttt{github.com/different-ai/openwork/blob/main/LICENSE}}
    \item \textbf{Issues \& Contributions:} \href{https://github.com/different-ai/openwork/issues}{\texttt{github.com/different-ai/openwork/issues}}
    \item \textbf{different-ai Organization:} \href{https://github.com/different-ai}{\texttt{github.com/different-ai}}
    \item \textbf{Skills Marketplace:} Available within the OpenWork desktop app under the \texttt{Skills} panel.
    \item \textbf{Build from Source:}
  \end{itemize}
\begin{lstlisting}
git clone https://github.com/different-ai/openwork
git submodule update --init --recursive
pnpm install && pnpm dev   # development
pnpm build                 # production
\end{lstlisting}
\end{frame}