%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
	
	\begin{center}
	{\Large Finding Your Persona \ldots

	Choosing Your Persona for Transition to Data Science}  
	\end{center}

\end{frame}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{The USER Persona}
        \begin{itemize}
            \item Leverage domain expertise.
            \item Utilize low/no code platforms like Weka or Knime.
            \item Build machine learning workflows easily.
            \item Ideal for managers, marketing professionals.
            \item Make data-driven decisions without extensive coding.
        \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{The DEVELOPER  Persona}
        \begin{itemize}
            \item Technical enthusiast with a programming flair.
            \item Expertise in scikit-learn, TensorFlow, PyTorch.
            \item Develop robust data science applications.
            \item Dive deep into machine learning techniques.
            \item Transform ideas into impact-ful solutions.
        \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{The RESEARCHER   Persona}
        \begin{itemize}
            \item Passionate about mathematics and innovation.
            \item Invent new techniques, contribute to research.
            \item Caters to deep R\&D professionals, PhD holders.
            \item Shape the data science landscape with created libraries.
            \item Be at the forefront of groundbreaking discoveries.
        \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Common to All}
        \begin{itemize}
            \item Solid foundation in machine and deep learning is crucial.
            \item Regardless of persona, these principles are essential.
            \item Embrace a continuous learning mindset.
            \item Stay updated with industry trends.
            \item Be open to expanding knowledge horizons.
        \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Choosing a Persona}
        \begin{itemize}
            \item Select a persona based on skills and interests.
            \item Allow projects and interests to guide your journey.
            \item Explore new areas of expertise.
            \item Expand your skill set accordingly.
            \item The right persona empowers excellence and lasting impact.
        \end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{So, Roles/Persona to chose from \ldots}
\begin{itemize}
\item Researcher: Advanced Mathematics, Programming, domain, Adv ML/DL, etc. You invent new things and write NeurIPS paper.
\item Implementation:
	\begin{itemize}
	\item Data Analyst: Bridge the gap between the data scientists and the business analysts, organizing and analyzing data to answer the questions the organization poses. 
	\item Data Engineer: Focus on developing, deploying, managing, and optimizing the organization’s data infrastructure and data pipelines.
	\item Data Scientist: Use/mine data, clean, build applications, reports, etc
	\end{itemize}
\item User level: Product Manager, Sales, Manager, etc
\end{itemize}
\end{frame}
