%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large Policies}

{\tiny (Ref:https://rasa.com/docs/rasa/core/policies/ )}
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{What are Policies?}
\begin{itemize}
\item Policy class decides which action to take at every step in the conversation
\item There are different policies to choose from, 
\item You can include multiple policies in a single Agent
\item Important hyperparameter in Policy is ``max\_history'' 
\item It controls how much dialogue history the model looks at to decide which action to take next.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{Config}
\begin{itemize}
\item config.yml file takes a policies key which you can use to customize the policies your assistant uses.
\item Below, the last two lines show how to use a custom policy class and pass arguments to it.
\end{itemize}

\begin{lstlisting}
policies:
  - name: "KerasPolicy"
    featurizer:
    - name: MaxHistoryTrackerFeaturizer
      max_history: 5
      state_featurizer:
        - name: BinarySingleStateFeaturizer
  - name: "MemoizationPolicy"
    max_history: 5
  - name: "FallbackPolicy"
    nlu_threshold: 0.4
    core_threshold: 0.3
    fallback_action_name: "my_fallback_action"
  - name: "path.to.your.policy.class"
    arg1: "..."
\end{lstlisting}



\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{Action Selection}
\begin{itemize}
\item At every turn, each policy defined in your configuration will predict a next action with a certain confidence level.
\item Final action is decided which has the highest confidence.
\item In case of clash, pre-defnied priorities are used.
\begin{itemize}
\item 5. FormPolicy
\item 4. FallbackPolicy and TwoStageFallbackPolicy
\item 3. MemoizationPolicy and AugmentedMemoizationPolicy
\item 2. MappingPolicy
\item 1. EmbeddingPolicy, KerasPolicy, and SklearnPolicy
\end{itemize}
\end{itemize}


\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{Keras Policy}
\begin{itemize}
\item Uses a neural network implemented in Keras to select the next action
\item The default architecture is based on an LSTM, 
\item But you can override the KerasPolicy.model\_architecture method to implement your own architecture.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{Mapping Policy}
\begin{itemize}
\item Used to directly map intents to actions.
\item The mappings are assigned by giving an intent the property triggers, e.g.:
\begin{lstlisting}
intents:
 - ask_is_bot:
     triggers: action_is_bot
\end{lstlisting}
\item An intent can only be mapped to at most one action. 
\item The bot will run the mapped action once it receives a message of the triggering intent. 
\item Afterwards, it will listen for the next message. 
\item With the next user message, normal prediction will resume.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{Memoization Policy}
\begin{itemize}
\item Just memorizes the conversations in your training data. 
\item It predicts the next action with confidence 1.0 if this exact conversation exists in the training data, otherwise it predicts None with confidence 0.0.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{Fallback Policy}
 The FallbackPolicy invokes a fallback action if at least one of the following occurs:
\begin{itemize}
\item The intent recognition has a confidence below nlu\_threshold.
\item The highest ranked intent differs in confidence with the second highest ranked intent by less than ambiguity\_threshold.
\item None of the dialogue policies predict an action with confidence higher than core\_threshold.
\end{itemize}

There is a special case called ``Two-Stage Fallback Policy'' also.
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile]\frametitle{Form Policy}
\begin{itemize}
\item The FormPolicy is an extension of the MemoizationPolicy which handles the filling of forms. 
\item Once a FormAction is called, the FormPolicy will continually predict the FormAction until all required slots in the form are filled. 
\end{itemize}
\end{frame}