%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large Graph Concepts} 
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large Graph Ontologies} 
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Key Ontology Standards}
    \begin{itemize}
        \item Ontology standards enable knowledge representation and reasoning.
        \item Used in AI, data management, and interoperability.
        \item Cover knowledge representation, upper-level ontologies, domain-specific models, reasoning frameworks, and enterprise standards.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Knowledge Representation Standards}
    \begin{itemize}
        \item \textbf{RDF}: W3C standard for structuring data using triples (subject-predicate-object).
        \item \textbf{OWL}: Expressive extension of RDF, enabling logical axioms and reasoning.
        \item \textbf{SPARQL}: Query language for RDF, supports complex queries.
        \item \textbf{SHACL}: Validates RDF graphs for schema enforcement.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Upper-Level Ontologies}
    \begin{itemize}
        \item \textbf{BFO}: Used in scientific, governmental, and military domains.
        \item \textbf{DOLCE}: Captures cognitive and linguistic ontological distinctions.
        \item \textbf{SUMO}: General-purpose ontology foundation for multiple domains.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Domain-Specific Ontologies}
    \begin{itemize}
        \item \textbf{UMLS}: Integrates multiple medical vocabularies.
        \item \textbf{FIBO}: Standardizes financial data models and transactions.
        \item \textbf{SNOMED CT}: Covers diseases, procedures, and anatomy.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Knowledge Graph and Reasoning Frameworks}
    \begin{itemize}
        \item \textbf{Protégé}: Open-source OWL and RDF ontology editor.
        \item \textbf{Neo4j}: Graph database with ontology plugins.
        \item \textbf{Stardog}: Commercial graph knowledge platform with reasoning.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Enterprise Ontology Standards}
    \begin{itemize}
        \item \textbf{ISO 15926}: Standard for process industries like oil and gas.
        \item \textbf{DCAT}: W3C vocabulary for describing data catalogs.
        \item \textbf{ISO/IEC 21838}: Defines top-level ontologies for interoperability.
    \end{itemize}
\end{frame}

