%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large Introduction}
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Introduction}
    \begin{itemize}
        \item Language models struggle with factual accuracy and real-world knowledge.
        \item Retrieval-Augmented Generation (RAG) improves accuracy using external text data.
        \item Traditional RAG has limitations in context understanding and scalability.
        \item GraphRAG leverages knowledge graphs for better retrieval and response generation.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Limitations of Traditional RAG}
    \begin{itemize}
        \item \textbf{Flat Retrieval:} Documents are treated as isolated entities.
        \item \textbf{Contextual Shortcomings:} Lacks deep semantic understanding.
        \item \textbf{Scalability Issues:} Slower retrieval with increasing data volume.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{What is GraphRAG?}
    \begin{itemize}
        \item Uses knowledge graphs instead of unstructured text.
        \item Captures entities, relationships, and hierarchical structures.
        \item Enables accurate, context-aware retrieval and response generation.
        \item Supports complex query handling with enhanced explainability.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Key Features of GraphRAG}
    \begin{itemize}
        \item \textbf{Structured Representation:} Uses knowledge graphs.
        \item \textbf{Contextual Retrieval:} Understands semantic relationships.
        \item \textbf{Efficient Processing:} Reduces computational cost.
        \item \textbf{Multi-Faceted Queries:} Synthesizes data from multiple sources.
        \item \textbf{Explainability:} More transparent than black-box models.
        \item \textbf{Continuous Learning:} Expands knowledge over time.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Applications of GraphRAG}
    \begin{itemize}
        \item \textbf{Healthcare:} Assists in diagnoses and treatment decisions.
        \item \textbf{Banking:} Detects fraudulent transactions using knowledge graphs.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{How GraphRAG Works}
    \begin{itemize}
        \item \textbf{Knowledge Graph Construction:} Extracts entities and relationships.
        \item \textbf{Knowledge Graph Summarization:} Generates hierarchical summaries.
        \item \textbf{Retrieval-Augmented Generation:} Uses local and global searches for queries.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Example of GraphRAG Representation}
    \begin{lstlisting}
    # Entities
    Type 2 Diabetes (Condition)
    High Blood Sugar Levels (Symptom)
    Nerve Damage (Complication)
    Kidney Disease (Complication)
    Cardiovascular Problems (Complication)

    # Relationships
    Type 2 Diabetes -> has_symptom -> High Blood Sugar Levels
    Type 2 Diabetes -> can_lead_to -> Nerve Damage
    Type 2 Diabetes -> can_lead_to -> Kidney Disease
    Type 2 Diabetes -> can_lead_to -> Cardiovascular Problems
    \end{lstlisting}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Advantages of GraphRAG}
    \begin{itemize}
        \item Structured knowledge representation.
        \item Context-aware and efficient retrieval.
        \item Handles complex queries effectively.
        \item Provides explainability and transparency.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Challenges of GraphRAG}
    \begin{itemize}
        \item \textbf{Complex Knowledge Graph Construction:} Requires sophisticated NLP techniques.
        \item \textbf{Data Dependency:} Performance relies on input data quality.
        \item \textbf{Scalability Issues:} Large graphs require significant computational resources.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Conclusion}
    \begin{itemize}
        \item GraphRAG enhances traditional RAG models using structured knowledge.
        \item Improves accuracy, context-awareness, and efficiency.
        \item Useful in various domains like healthcare and banking.
        \item A promising approach for future AI-powered knowledge retrieval.
    \end{itemize}
\end{frame}
