%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{References}

\begin{itemize}
\item What is Geometric Deep Learning? - Flawnson Tong
\item Towards Geometric Deep Learning I-V + Conclusions - Michael Bronstein
\item AMMI Course "Geometric Deep Learning" 2022 - Michael Bronstein
\item Shape Analysis (Spring 2023) -  Justin Solomon, MIT
\item Equivariant Neural Networks - DeepFindr, wonderful info on background of symmetries, groups and GNNs
\item Stanford Graph ML Projects - https://medium.com/stanford-cs224w
\item GDL website https://geometricdeeplearning.com/
\item Book: Geometric Deep Learning: Grids, Groups, Graphs, Geodesics, and Gauges - Michael M. Bronstein, Joan Bruna, Taco Cohen, Petar Velickovic
\item GAT - Graph Attention Network (PyTorch) https://github.com/gordicaleksa/pytorch-GAT
\item Graph Deep Learning Lab - Xavier Bresson https://graphdeeplearning.github.io/authors/xavier-bresson/
\item Ekagra Ranjan : https://ekagra-ranjan.github.io/
\item Charu Sharma : https://www.iiit.ac.in/people/faculty/Charusharma/
\item Tanmoy Chakraborty : https://www.tanmoychak.com/
\end{itemize}
	  
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Further Reading}

\begin{block}{Foundational Papers}
\begin{itemize}
\item Bronstein et al. (2021). "Geometric Deep Learning: Grids, Groups, Graphs, Geodesics, and Gauges." arXiv:2104.13478
\item Cohen \& Welling (2016). "Group Equivariant Convolutional Networks." ICML
\item Kipf \& Welling (2017). "Semi-Supervised Classification with Graph Convolutional Networks." ICLR
\item Thomas et al. (2018). "Tensor field networks: Rotation- and translation-equivariant neural networks for 3D point clouds." arXiv:1802.08219
\end{itemize}
\end{block}

\begin{block}{Books and Surveys}
\begin{itemize}
\item Hamilton (2020). "Graph Representation Learning." Morgan \& Claypool
\item Wu et al. (2021). "A Comprehensive Survey on Graph Neural Networks." IEEE TNNLS
\item Bekkers (2020). "B-Spline CNNs on Lie Groups." ICLR
\end{itemize}
\end{block}

\begin{block}{Online Resources}
\begin{itemize}
\item Geometric Deep Learning website: \texttt{geometricdeeplearning.com}
\item PyTorch Geometric tutorials: \texttt{pytorch-geometric.readthedocs.io}
\item Stanford CS224W: Machine Learning with Graphs
\item Petar Veličković's blog on Geometric Deep Learning
\end{itemize}
\end{block}

\end{frame}