%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large Certifications}
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}\frametitle{Current Certifications}
\begin{itemize}
\item  ``Neo4j Certified Professional'' (80 questions, 60 minutes) Foundation
% \item [Check if available] ``Neo4j 4.0 Certification'' (30 questions, 45 minutes), Specific important features of the version 4
\item ``Neo4j Graph Data Science Certification'' (40 questions, 60 minutes)
\end{itemize}


\begin{itemize}
\item All are free, multiple times (gap of 24 hours, 80\% passing)
\item All exam questions are either single-selection or multiple-choices, with four choices each
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}\frametitle{Neo4j Certified Professional}

\begin{itemize}
\item Charges: No cost – it’s completely free!
\item Must score 80\% or above in order to be certified.
\item Can retake the certification exam later, after a cooling off period of 24 hours.
\item Prerequisite Courses:
	\begin{itemize}
	\item ``Neo4j Fundamentals''
	\item ``Cypher Fundamentals''
	\item ``Graph Data Modeling Fundamentals''
	\item ``Importing CSV Data into Neo4j''
	\item `Intermediate Cypher Queries''
	\item ``Building Neo4j Applications with *''
	\end{itemize}
\end{itemize}
\end{frame}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}\frametitle{Neo4j Graph Data Science  Certification}
\begin{itemize}
\item Charges: No cost – it’s completely free!
\item Must score 80\% or above in order to be certified.
\item Can retake the certification exam later, after a cooling off period of 24 hours.
\item Prerequisite Courses:
	\begin{itemize}
	\item ``Neo4j Fundamentals''
	\item ``Cypher Fundamentals''
	\item `Intermediate Cypher Queries''
	\item ``Building Neo4j Applications with *''
	\item ``Introduction to Neo4j Graph Data Science''
	\item ``Neo4j Graph Data Science Fundamentals''
	\end{itemize}
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}\frametitle{Tips}
\begin{itemize}
\item This is not a user test… it’s a user and DBA test.
\item  Pay attention to Cypher queries. The devil’s in the details.
\item Develop query equivalence to second nature.
\item The test is designed on the basis of the Enterprise edition (not 'Community' edition)
\item The test is generally well-written (so, no goof ups)
\item TRUNCATE BRAINSPACE sql\_clauses;
\item Have a 30,000ft overview of the subject
\item Algorithms, statistics and aggregation
\item Practice on Northwind DB but know the Movie DB like the back of your hand.
\item  Ref cards are your friends.
\end{itemize}

{\tiny (Ref:10 TIPS FOR PASSING THE NEO4J CERTIFIED PROFESSIONAL EXAMINATION
- CHRIS VON CSEFALVAY )}

\end{frame}


% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{frame}\frametitle{Neo4j 4.0 Certification}
% Prerequisites:

% \begin{itemize}
% \item Overview of Neo4j 4.x
% \item Querying with Cypher in Neo4j 4.x
% \item Creating Nodes and Relationships in Neo4j 4.x
% \item Using Indexes and Query Best Practices in Neo4j 4.x
% \item Importing Data with Neo4j 4.x
% \item Basic Neo4j 4.x Administration
% \end{itemize}

% Distribution:

% \begin{itemize}
% \item RBAC – 10 questions (32\%)
% \item MultiDB – 8 questions (24\%)
% \item Cypher – 6 questions (20\%)
% \item General – 4 questions (16\%)
% \item Fabric – 2 questions (8\%)
% \end{itemize}
% \end{frame}

