
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large Applications of Prompts to Marketing}
\end{center}
\end{frame}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{LinkedIn Post Generator}

{\tiny (Ref: A Beginner's Guide to Prompt Engineering with ChatGPT - Datacamp)}


\begin{lstlisting}
Act as a social media copywriter. Create a Linkedin post about a
webinar titled "Generative Al Applications for Enterprise Organizations:
Unlocking the Power of Creative Automation". Follow my instructions
and follow them very carefully! Do not deviate away from my
instructions!

1. Provide a step by step reasoning of why you're making the decisions
you are making
2. Write an engaging, professional LinkedIn post.
3. Make your Linkedin post around 800 characters.
4. Use small paragraphs, add lists and line breaks, a small amount of
emojis, to make the post easier to read.
5. Have a credible, professional, and accessible voice and tone. A
good example would be that of DataCamp, McKinsey, and other
reputable organizations.
6. Avoid bad copy.
7. Avoid inaccuracies.
\end{lstlisting}		
		

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Twitter Thread Creator}

{\tiny (Ref: A Beginner's Guide to Prompt Engineering with ChatGPT - Datacamp)}


\begin{lstlisting}
Act as a social media copywriter. Create a Twitter thread about a
webinar titled "Generative Al Applications for Enterprise Organizations:
Unlocking the Power of Creative Automation". Follow my instructions.
and follow them very carefully! Do not deviate away from my
instructions!

1. Provide a step by step reasoning of why you're making the decisions
you are making
2. Write an engaging twitter thread.
3. Make your twitter thread post around 'n' tweets long
4. Make your twitter thread informative, engaging, and visually
appealing by using one emoji per tweet.
5. Have a credible, professional, and accessible voice and tone. A
good example would be that of DataComp, McKinsey, and other reputable organizations.
6. Avoid bad copy.
7. Avoid inaccuracies
\end{lstlisting}		
		

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Ad Copy Creator}

{\tiny (Ref: A Beginner's Guide to Prompt Engineering with ChatGPT - Datacamp)}


\begin{lstlisting}
Act as a paid media copywriter. Create {Social Channel} paid ad copy about a
webinar titled "Generative Al Applications for Enterprise Organizations:
Unlocking the Power of Creative Automation". Follow my instructions.
and follow them very carefully! Do not deviate away from my
instructions!

1. Provide a step by step reasoning of why you're making the decisions
you are making
2. Write an engaging {platform} ad, that will maximize conversions.
3. Make your {platform} post is around {character limit} long
4. Have a credible, professional, and accessible voice and tone. A
good example would be that of DataComp, McKinsey, and other reputable organizations.
6. Avoid bad copy.
7. Avoid inaccuracies
8. Provide at least 3 variations
\end{lstlisting}		
		
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Email Generator}

{\tiny (Ref: A Beginner's Guide to Prompt Engineering with ChatGPT - Datacamp)}


\begin{lstlisting}
Act as a sales person. Create a cold email to a prospect discussing
{service they might be interested in} and how it can be useful for their
organizations. Follow my instructions and follow them very carefully!
Do not deviate away from my instructions!

1. Provide a step by step reasoning of why you'e making the decisions
you are making
2 Make sure the email is professional, personable, concise, and
useful
3. Make sure to speak to their pain point, as {title}, the prospect
usually has {x} pain point
4. Avoid bad copy.
5. Be respectful of their time, do not add any useless details about
how you prepared for writing this email
6. Avoid inaccuracies.
7. Make sure the email is at most 100 words long
\end{lstlisting}		
		
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Contract Reviewer}

{\tiny (Ref: A Beginner's Guide to Prompt Engineering with ChatGPT - Datacamp)}


\begin{lstlisting}
Act as a lawyer. Explain the following legal language to me. Follow my
instructions and follow them very carefully.

1. Provide a step by step reasoning of why you're making the decisions
you are making
2 Explain the legal language to me as a layman. | am not a lawyer.
3. Avoid inaccuracies.

"Inspection of Goods & Rejection. The Buyer is entitled to inspect the
Goods upon delivery. if the Goods are unacceptable for any reason,
the Buyer must reject them at the time of delivery or within five (5)
business days from the date of delivery. If the Buyer has not rejected
the Goods within five (5) business days from the date of delivery, the
Buyer shall have waived any right to reject that specific delivery of
Goods. In the event the Buyer rejects the Goods, the Buyer shall allow
the Seller a reasonable time to cure the deficiency. A reasonable time
period shall be determined by industry standards for the particular goods
as well as the Seller and the Buyer.
\end{lstlisting}		
		
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Customer Support Problem Solver}

{\tiny (Ref: A Beginner's Guide to Prompt Engineering with ChatGPT - Datacamp)}


\begin{lstlisting}
Act as a customer service representative. Please respond to the
following customer ticket. Follow my instructions and follow them very
carefully! Do not deviate away from my instructions!

1. Provide a step by step reasoning of why you're making the decisions
you are making
2 Make sure your response is comprehensive, concise, clear, and
cordial
3. Provide the customers with the answer outlined below.
4. Avoid inaccuracies.
5. Set a word count limit of 50 words long

- Customer Problem Example used: Will there be a recording of the
webinar available?
- Solution: Yes, webinars and recordings will be sent to you via email within a few days.

\end{lstlisting}		
		
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Creating Social Media Posts}
\begin{itemize}
\item Basic Prompt: "Write a tweet promoting our new product launch in 280 characters or less."
\item Comprehensive Prompt
	\begin{itemize}
	\item Context: You are a seasoned Marketing professional working for a well-established company in the tech industry. Your company has recently launched a new product and is looking to generate buzz and engage with its target audience through social media platforms. As the lead marketer, you are responsible for creating compelling social media posts that effectively promote the product, drive user engagement, and ultimately contribute to increasing sales and brand awareness.
	\item Prompt: Create a series of five social media posts (one for each major platform) that effectively promote the company's new product. Each post should be concise, visually appealing, and tailored to the specific platform's requirements and audience preferences. Your goal is to capture the attention of the target audience, spark their interest, and encourage them to take action (e.g., visit the company's website, sign up for a newsletter, or make a purchase).
	\item Guardrails: 
		\begin{itemize}
		\item The total character count for each social media post should not exceed 280 characters for Twitter, 2,200 characters for LinkedIn, and 2,200 characters for Facebook.
		\item Each post should include relevant hashtags, mentions, or links, as appropriate, to maximize visibility and encourage social sharing.
		\item Ensure that the language and tone of the posts align with the company's brand voice and values.
		\item Utilize compelling visuals, such as images or videos, to enhance the impact of each post.
		\item Tailor the content and format of the posts to suit the specific platform's requirements and audience preferences. For example, use hashtags and concise messaging for Twitter, a more professional tone for LinkedIn, and engaging storytelling for Facebook.
		\end{itemize}	
	\end{itemize}
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Creating Social Media Posts}
Comprehensive Prompt
{\tiny
\begin{itemize}
\item Context: You are a seasoned Marketing professional working for a well-established company in the tech industry. Your company has recently launched a new product and is looking to generate buzz and engage with its target audience through social media platforms. As the lead marketer, you are responsible for creating compelling social media posts that effectively promote the product, drive user engagement, and ultimately contribute to increasing sales and brand awareness.
\item Prompt: Create a series of five social media posts (one for each major platform) that effectively promote the company's new product. Each post should be concise, visually appealing, and tailored to the specific platform's requirements and audience preferences. Your goal is to capture the attention of the target audience, spark their interest, and encourage them to take action (e.g., visit the company's website, sign up for a newsletter, or make a purchase).
\item Guardrails: 
	\begin{itemize}
	\item The total character count for each social media post should not exceed 280 characters for Twitter, 2,200 characters for LinkedIn, and 2,200 characters for Facebook.
	\item Each post should include relevant hashtags, mentions, or links, as appropriate, to maximize visibility and encourage social sharing.
	\item Ensure that the language and tone of the posts align with the company's brand voice and values.
	\item Utilize compelling visuals, such as images or videos, to enhance the impact of each post.
	\item Tailor the content and format of the posts to suit the specific platform's requirements and audience preferences. For example, use hashtags and concise messaging for Twitter, a more professional tone for LinkedIn, and engaging storytelling for Facebook.
	\end{itemize}	
\end{itemize}
}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Example Social Media Posts}
{\tiny
\begin{itemize}
\item Twitter Post (280 characters max):
"Introducing our game-changing product that will revolutionize your life! Discover how our [Product Name] boosts productivity, saves time, and amplifies results. Click the link below to learn more! \#ProductLaunch \#TechInnovation \#BoostYourLife"
\item LinkedIn Post (2,200 characters max):
"Attention professionals! We're excited to announce our latest breakthrough – [Product Name]. Designed to streamline workflows and optimize efficiency, it's the perfect tool to elevate your productivity game. Join the revolution today! [Include compelling visual and link] \#ProductivityTools \#WorkSmarter \#LinkedInExclusive"
\item Facebook Post (2,200 characters max):
"Attention all dreamers and doers! We're thrilled to introduce [Product Name], the ultimate solution to help you achieve your goals. Dive into the possibilities, unleash your potential, and embark on a journey of success with our game-changing product. [Include captivating video and call-to-action] \#DreamBig \#SuccessStories \#AchieveMore"
\item Instagram Post (2,200 characters max):
"Ready to level up your game? Introducing [Product Name], the game-changer you've been waiting for! Take charge of your destiny and unlock new possibilities with our revolutionary tool. Double tap if you're excited to learn more! [Include eye-catching image and swipe-up link] \#GameChanger \#UnlockPotential \#InstaInnovation"
\item TikTok Post (280 characters max):
"Brace yourself for a mind-blowing experience! Get ready to witness the future of [Product Category] with our game-changing [Product Name]. Join the trendsetters, unlock your potential, and transform your life with this revolutionary tool! \#ProductLaunch \#TikTokTrendsetters \#GameChanger"
\end{itemize}
}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Writing Email Subject Lines}
\begin{itemize}
\item Basic Prompt: "Craft a subject line for our upcoming newsletter that will entice subscribers to open it."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Generating Blog Post Titles}
\begin{itemize}
\item Basic Prompt: "Come up with a catchy title for a blog post about the benefits of our product."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Creating Ad Copy}
\begin{itemize}
\item Basic Prompt: "Write a Facebook ad that will encourage users to sign up for our free trial."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Developing Website Copy}
\begin{itemize}
\item Basic Prompt: "Write a landing page headline that will grab the attention of visitors and encourage them to learn more about our product."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Crafting Product Descriptions}
\begin{itemize}
\item Basic Prompt: "Write a product description that highlights the key features and benefits of our new release."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Creating Video Scripts}
\begin{itemize}
\item Basic Prompt: "Write a script for a product explainer video that will engage viewers and encourage them to learn more about our offering."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Developing Podcast Episode Titles}
\begin{itemize}
\item Basic Prompt: "Come up with a title for our next podcast episode that will entice listeners to tune in."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Writing Press Releases}
\begin{itemize}
\item Basic Prompt: "Craft a press release announcing our latest funding round that will grab the attention of journalists and investors."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Developing Case Studies}
\begin{itemize}
\item Basic Prompt: "Write a case study that showcases how our product helped a customer achieve their goals."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Creating Infographics}
\begin{itemize}
\item Basic Prompt: "Develop an infographic that illustrates the key benefits of our product in a visually engaging way."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Crafting Social Media Ad Copy}
\begin{itemize}
\item Basic Prompt: "Write a Twitter ad that will encourage users to click through to our landing page and learn more about our product."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Developing Email Body Copy}
\begin{itemize}
\item Basic Prompt: "Write the body copy for an email promoting our upcoming webinar that will encourage recipients to register."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Creating Website Headlines}
\begin{itemize}
\item Basic Prompt: "Write a headline for our homepage that will grab the attention of visitors and encourage them to explore our site."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Developing Product Demo Scripts}
\begin{itemize}
\item Basic Prompt: "Write a script for a product demo video that will showcase the key features and benefits of our offering."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Crafting Social Media Captions}
\begin{itemize}
\item Basic Prompt: "Write an Instagram caption that will encourage users to engage with our brand and share our content."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Developing White Paper Titles}
\begin{itemize}
\item Basic Prompt: "Come up with a title for our latest white paper that will entice readers to download and read it."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Creating Webinar Titles}
\begin{itemize}
\item Basic Prompt: "Come up with a title for our upcoming webinar that will encourage attendees to register and tune in."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Developing Email Subject Lines for Abandoned Cart Emails}
\begin{itemize}
\item Basic Prompt: "Craft a subject line for an abandoned cart email that will encourage customers to complete their purchase."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Creating Social Media Graphics}
\begin{itemize}
\item Basic Prompt: "Develop a social media graphic that highlights the key benefits of our product in a visually engaging way."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Crafting Ad Headlines}
\begin{itemize}
\item Basic Prompt: "Write a headline for a Google ad that will encourage users to click through to our landing page and learn more about our product."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Developing Email Body Copy for Welcome Emails}
\begin{itemize}
\item Basic Prompt: "Write the body copy for a welcome email that will encourage new subscribers to engage with our brand and learn more about our offerings."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Creating Social Media Videos}
\begin{itemize}
\item Basic Prompt: "Develop a short video for Instagram that showcases the key features and benefits of our product."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Crafting Ad Descriptions}
\begin{itemize}
\item Basic Prompt: "Write a description for a Facebook ad that will encourage users to click through to our landing page and learn more about our product."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Developing Email Subject Lines for Promotional Emails}
\begin{itemize}
\item Basic Prompt: "Craft a subject line for a promotional email that will encourage recipients to take advantage of our limited-time offer."
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{More Prompts}
{\tiny
\begin{itemize}
\item Suggest some effective marketing strategies for a [startup / business / product / company]
\item How to improve social media engagement and increase followers for [product / business]?
\item What are some creative ways to promote [product or service name] on social media?
\item What are the most important SEO techniques for improving website traffic?
\item What are some effective strategies for improving the SEO of a [type of website]?
\item Can you help me come up with a creative advertising campaign for a new [product name] launch?
\item How can I analyze and interpret data from my marketing campaigns to make better decisions?
\item What are some best practices for email marketing to increase open rates and click-through rates?
\item How can I leverage influencer marketing to promote my [product / service name / brand]?
\item Can you provide some tips for creating a successful content marketing strategy?
\item What are the most effective ways to convert website visitors into customers?
\item How can I optimize my landing pages for better conversion rates?
\end{itemize}
}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{More Prompts}
{\tiny 
\begin{itemize}
\item Suggest a catchy slogan for [product or service name]
\item Create a list of potential target audiences for [product or service name]
\item Develop a content marketing plan for [product or service name] that includes blog posts, social media, and email campaigns
\item What are some proven methods for building brand awareness for [company / product name] among a wider audience?
\item Create a buyer persona for [product or service name] that includes their demographics, interests, and pain points
\item What are some effective ways to measure the ROI of [marketing campaign or strategy]?
\item Develop a lead nurturing email campaign for [product or service name] that encourages prospects to take action
\item Provide some ideas for blog posts about [topic]
\item Write a description for [product or service or company]
\item Suggest inexpensive ways to promote [product or service or company] without social media
\item How to obtain high quality backlinks to improve [website] SEO
\end{itemize}
}
\end{frame}