
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large Applications of Prompts to Marketing}
\end{center}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Mathematics Prompts}
    \begin{itemize}
        \item Explain the concept of fractions in your own words.
        \item Solve a multi-step word problem involving percentages.
        \item Describe the relationship between the area and perimeter of a rectangle.
        \item Illustrate the difference between rational and irrational numbers using examples.
        \item Create a step-by-step guide to solving linear equations.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Science Prompts}
    \begin{itemize}
        \item Explain the process of photosynthesis in a way a middle school student can understand.
        \item Create a diagram to illustrate the structure of an animal cell.
        \item Describe the different types of energy and provide examples of each.
        \item Conduct a simple experiment to demonstrate the properties of matter.
        \item Explain the role of the human respiratory system and its importance.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Language Arts Prompts}
    \begin{itemize}
        \item Write a short story about a middle school student who discovers a hidden talent.
        \item Analyze the theme of courage in a literary work of your choice.
        \item Create a persuasive essay on the importance of reading for personal growth.
        \item Describe the characteristics of different literary genres (e.g., poetry, fiction, non-fiction).
        \item Retell a classic fairy tale or folktale from a different perspective.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Social Studies Prompts}
    \begin{itemize}
        \item Describe the key events and consequences of the American Revolution.
        \item Analyze the impact of a significant historical figure on their society.
        \item Create a timeline of major technological advancements in the 20th century.
        \item Explain the role of the United Nations in promoting global cooperation.
        \item Investigate the cultural traditions and customs of a specific region or country.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Creative Arts Prompts}
    \begin{itemize}
        \item Design a poster or brochure promoting a school event or activity.
        \item Compose a short musical piece using a specific set of notes or instruments.
        \item Create a three-dimensional model of a famous landmark or building.
        \item Write and perform a short skit or play with a middle school-appropriate theme.
        \item Illustrate a scene from a book or story using various art mediums.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{General Learning Prompts}
    \begin{itemize}
        \item Explain the importance of setting personal goals and developing a growth mindset.
        \item Describe strategies for effective time management and organization.
        \item Create a step-by-step guide to conducting research on a topic of your choice.
        \item Analyze the benefits of collaborative learning and teamwork.
        \item Describe the role of technology in enhancing the learning experience.
    \end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Chain of Thoughts Prompts}
    \begin{itemize}
        \item Explain the steps to solve a multi-step algebra problem, showing your work.
        \item Describe the thought process you would use to research and write a biography on a famous scientist.
        \item Walk through the decision-making steps a middle school student would take to plan a class field trip.
        \item Demonstrate the logical reasoning you would use to argue in favor of implementing sustainable practices in your school.
        \item Outline the mental strategies you would employ to memorize the key events of the Civil War.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Project Work Prompts}
    \begin{itemize}
        \item Design and build a model of a renewable energy source, such as a small-scale wind turbine or solar panel.
        \item Create a group project to research and present on a historical event, highlighting its causes and consequences.
        \item Develop a plan for a school-wide recycling program, including a proposal and implementation strategy.
        \item Organize a community service project, such as a neighborhood clean-up or a food drive, and document the process.
        \item Construct a scale model of the solar system, demonstrating the relative sizes and distances of the planets.
    \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Essay Writing Prompts}
    \begin{itemize}
        \item Write a persuasive essay arguing for the importance of physical education in middle schools.
        \item Compose a personal narrative describing a significant moment that shaped your academic or personal growth.
        \item Analyze the theme of friendship in a novel or short story, and discuss its relevance to middle school students.
        \item Describe the impact of technology on the way you learn and communicate, and how it has affected your daily life.
        \item Write an expository essay explaining the scientific principles behind a natural phenomenon, such as the formation of rainbows.
    \end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Ethical Considerations}
    \begin{itemize}
        \item Ensure the content is age-appropriate and does not expose middle school students to harmful or inappropriate material.
        \item Promote inclusivity and respect for diversity, avoiding biased or discriminatory language or content.
        \item Encourage critical thinking and problem-solving over rote memorization or simply seeking the "right" answer.
        \item Protect student privacy and data, and avoid prompts that require the sharing of personal information.
        \item Emphasize the responsible and ethical use of technology, such as avoiding plagiarism or misuse of AI-generated content.
    \end{itemize}
\end{frame}