%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large Introduction}
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Autonomous AI Agents}
  \begin{itemize}
    \item Remarkable advancement in artificial intelligence.
    \item Combined power of multiple ChatGPTs collaboratively engaging in dynamic conversations.
    \item Collaborative approach yields astonishing enhancements in performance and capabilities.
    \item Contrasted with using a single AI, such as ChatGPT, in isolation.
    \item Ability to assume distinct roles within a team.
    \item Similar to professionals in various fields.
    \item Roles include project managers, developers, designers, etc.
    \item Each agent contributes specialized expertise to the conversation.
    \item Multiplicity of roles enables addressing broader challenges.
    \item Provides a holistic and efficient solution-seeking experience.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{The Blueprint}
  \begin{itemize}
    \item Planning: Reflects on past experiences, offers self-critiques, and breaks down tasks into manageable steps using sub-goal decomposition.
    \item Memory: Utilizes sensory, short-term, and long-term memory for real-time data processing, task-specific information, and retaining knowledge/experiences.
    \item Tools: Equipped with a virtual toolbox, accessing calendars, calculators, search engines, and other resources for versatile problem-solving.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Flow: The Symphony}
  \begin{itemize}
    \item Task Decomposition: Breaks down tasks into smaller, more manageable components for enhanced efficiency and accuracy.
    \item Model Selection: Chooses the most suitable Large Language Model (LLM) for the task to align actions with desired outcomes.
    \item Task Execution: Executes tasks with precision and speed, leveraging planning, memory, and tools.
    \item Response Generation: Generates contextually relevant and accurate responses, be it drafting a report, answering questions, or making decisions.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Challenges in LLM-Centered Agents}
  \begin{itemize}

    \item Finite Context Length: Restricted context capacity limits inclusion of historical information, detailed instructions, API call context, and responses.
    \item System design works with limited communication bandwidth, impacting mechanisms like self-reflection.
    \item Challenges in long-term planning and task decomposition: LLMs struggle to adjust plans when faced with unexpected errors.
    \item Planning over lengthy history and exploring solution space remain challenging for LLMs.
    \item Less robust compared to humans who learn from trial and error.
    \item Reliability of Natural Language Interface: Current agent system relies on natural language as an interface, but the reliability of model outputs is questionable.
    \item LLMs may make formatting errors and occasionally exhibit rebellious behavior (e.g., refuse to follow an instruction).
    \item Agent demo code often focuses on parsing model output due to these reliability issues.
  \end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Real-World Applications}

  \begin{itemize}
    \item BabyAGI: Pioneering AI learning system.
    \item AutoGPT: Automates content generation.
    \item GPT Engineer: Assists in coding and software development.
    \item AAA-powered entities pushing AI boundaries in various industries.
  \end{itemize}
\end{frame}

