\input{template_homework}



\newcommand{\hmwkTitle}{Machine Learning For Mechanical Engineering}
\newcommand{\hmwkDueDate}{February 20, 2017}
\newcommand{\hmwkClass}{Test 1 (T1)}
\newcommand{\hmwkClassTime}{TY Mechanical Engineering}
\newcommand{\hmwkClassInstructor}{CoEP}
\newcommand{\hmwkAuthorName}{\textbf{Yogesh Kulkarni}}

\title{
%    \vspace{2in}
    \textmd{\textbf{\hmwkClass\\\ \hmwkTitle}}\\
    {Due\ on\ \hmwkDueDate\ at 5pm}\\
    \vspace{0.1in}\large{{\hmwkClassInstructor\ \hmwkClassTime}}
}

\author{\hmwkAuthorName}
\date{}

\begin{document}

\maketitle

%\pagebreak

\begin{homeworkProblem}
(Marks: 10)
\begin{itemize}
\item Pick up any Mechanical Engineering related numeric problem, such as Geometric Transformations, Finite Element Truss calculations, etc.
\item Write Python program to solve the calculations.
\item Ask user to provide the inputs needed.
\item Present the results to the user, preferably graphically.
\item Grading will be done on the complexity of the problem chosen and the usage of various features of Python programming.
\item Problem chosen by each student should be unique. Duplicate problems may not fetch marks.
\end{itemize}
\end{homeworkProblem}

\begin{homeworkProblem}
(Marks: 10)
\begin{itemize}
\item Pick up any dataset from UCI or Kaggle, such as Titanic Disaster dataset, which has missing values.
\item Write Python program for data preparation.
\item Read the dataset in Pandas and do following steps:
\begin{itemize}
\item Summary statistics of relevant columns.
\item Conversion of categorical variables to numbers.
\item Standardization/normalization of relevant columns.
\item Crosstab and Correlation plots analysis.
\item Formulation of new features is encouraged.
\item Imputation of Missing values.
\item Output Cross Validation (4) matrices from training file.
\end{itemize}
\item Grading will be done on the usage of various features of Pandas.
\end{itemize}
\end{homeworkProblem}


\end{document}