\input{template_homework}



\newcommand{\hmwkTitle}{Machine Learning For Mechanical Engineering}
\newcommand{\hmwkDueDate}{April 2, 2018}
\newcommand{\hmwkClass}{Test 2 (T2)}
\newcommand{\hmwkClassTime}{TY Mechanical Engineering}
\newcommand{\hmwkClassInstructor}{CoEP}
\newcommand{\hmwkAuthorName}{\textbf{Yogesh Kulkarni}}

\title{
%    \vspace{2in}
    \textmd{\textbf{\hmwkClass\\\ \hmwkTitle}}\\
    {Due\ on\ \hmwkDueDate\ at 5pm}\\
    \vspace{0.1in}\large{{\hmwkClassInstructor\ \hmwkClassTime}}
}

\author{\hmwkAuthorName}
\date{}

\begin{document}

\maketitle

\section{Common Instructions}
\begin{itemize}
\item Pick up any suitable dataset from UCI or Kaggle, etc, preferably related to Mechanical Engineering, Manufacturing, etc.  If not available, other domains are ok as well.
\item Each student should have a unique dataset, eg for 30 students, $30x2$ unique datasets.
\item Write IPython Notebooking to solve the problem with steps as follows:
\begin{itemize}
\item Read the dataset in Pandas
\item Summary statistics of relevant columns.
\item Conversion of categorical variables to numbers.
\item Standardization/normalization of relevant columns.
\item Correlation plots analysis.
\item Formulation of new features, if needed.
\item Imputation of Missing values, if any.
\item Cross Validation (4) matrices from training file.
\item Solve (fit, predict, score) using the \textbf{suggested model} from Scikit Learn.
\item Print \textbf{appropriate evaluation metrics}, plotting would be preferred if possible.
\end{itemize}
\item Grading will be done on the complexity of the problem chosen.
\end{itemize}

\begin{homeworkProblem}
\textbf{Regression}
(Marks: 10)
\begin{itemize}
\item Model: Linear Regression
\item Evaluation: $R^2$
\end{itemize}
\end{homeworkProblem}

\begin{homeworkProblem}
\textbf{Classification}
(Marks: 10)
\begin{itemize}
\item Model: Logistic Regression or Decision Tree
\item Evaluation: $F1$
\end{itemize}
\end{homeworkProblem}


\end{document}