%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large Linear Algebra Summary}

{\tiny (Ref: A Gentle Introduction to Linear Algebra
 - Json Brownlee)}

\end{center}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile] \frametitle{Summary}

\begin{itemize}

\item Linear algebra is about linear combinations.
\item Linear algebra is the study of lines and planes, vector spaces and mappings that are required for linear transforms
\item Applications of Linear Algebra
\begin{itemize}
\item Matrices in Engineering, such as a line of springs.
\item Graphs and Networks, such as analyzing networks.
\item Computer Graphics, such as the various translation, rescaling and rotation of images.
\end{itemize}

\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frame}[fragile] \frametitle{Further Reading}

Books:

\begin{itemize}
\item Introduction to Linear Algebra by Serge Lang
\item Introduction to Linear Algebra, Gilbert Strang, 2016.
\item  Numerical Linear Algebra, Lloyd N. Trefethen, 1997.
\item  Linear Algebra and Matrix Analysis for Statistics, Sudipto Banerjee, Anindya Roy, 2014.
\end{itemize}

Courses:

\begin{itemize}

\item  “Linear Algebra for machine learning” - Patrick van der Smagt 
\item “Machine Learning – 03. Linear Algebra Review“. Playlist at Youtube
\item  Linear Algebra stream on Khan Academy.
\end{itemize}



\end{frame}
