%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large Applications to Agricultural Engineering}
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Introduction}
  \begin{itemize}
    \item \textbf{Definition}: Machine Learning (ML) uses algorithms to enable systems to learn and make decisions from data.
    \item \textbf{Significance in Agriculture}:
      \begin{itemize}
        \item Enhancing crop yields and reducing waste.
        \item Optimizing resource usage (water, fertilizers).
      \end{itemize}
    \item \textbf{Why Now?}:
      \begin{itemize}
        \item Increased data availability (IoT, sensors, satellites).
        \item Advances in computing power and ML algorithms.
      \end{itemize}
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Core Concepts}
  \begin{itemize}
    \item \textbf{Supervised Learning}:
      \begin{itemize}
        \item Models learn from labeled data.
        \item Examples: Crop yield prediction, pest detection.
      \end{itemize}
    \item \textbf{Unsupervised Learning}:
      \begin{itemize}
        \item Models identify patterns in unlabeled data.
        \item Examples: Soil classification, weather clustering.
      \end{itemize}
    \item \textbf{Reinforcement Learning}:
      \begin{itemize}
        \item Models learn optimal actions via rewards.
        \item Examples: Irrigation management.
      \end{itemize}
  \end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Application Use-Cases}
  \begin{itemize}
    \item \textbf{Precision Farming}:Crop monitoring using drones and sensors.
    \item \textbf{Predictive Analytics}: Weather forecasting, yield prediction.
    \item \textbf{Livestock Management}: Health monitoring and feeding optimization.
    \item \textbf{Pest and Disease Detection}: Image recognition for early detection.
  \end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Examples of ML Algorithms}
  \begin{itemize}
    \item \textbf{Regression}: Predicting crop yields based on weather.
    \item \textbf{Classification}: Identifying healthy vs. diseased plants.
    \item \textbf{Clustering}: Grouping soil types for specific crops.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Regression: Yield Prediction}
	\begin{itemize}
	\item Model: Linear Regression
	\item Data: Historical weather data, soil data, fertilizer inputs
	\item Code Snippet (Python with scikit-learn):
	\end{itemize}
	
	\begin{lstlisting}
	from sklearn.linear_model import LinearRegression
	model = LinearRegression()
	model.fit(X_train, y_train)
	y_pred = model.predict(X_test) 1 
	\end{lstlisting}	
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Classification: Disease Detection}
	\begin{itemize}
	\item Model: Support Vector Machine (SVM)
	\item Data: Image data of plant leaves, spectral data
	\item Code Snippet (Python with scikit-learn):
	\end{itemize}
	
\begin{lstlisting}
from sklearn.svm import SVC
model = SVC(kernel='linear')
model.fit(X_train, y_train)
y_pred = model.predict(X_test)
\end{lstlisting}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Clustering: Soil Type Identification}
	\begin{itemize}
	\item Model: K-Means Clustering
	\item Data: Soil pH, nutrient levels, texture data
	\item Code Snippet (Python with scikit-learn):
	\end{itemize}

\begin{lstlisting}
from sklearn.cluster import KMeans
kmeans = KMeans(n_clusters=3)
kmeans.fit(X)
labels = kmeans.labels_
\end{lstlisting}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Conclusions}
  \begin{itemize}
    \item \textbf{Machine Learning in Agriculture}:
      \begin{itemize}
        \item Offers innovative solutions to complex problems.
        \item Drives efficiency, productivity, and sustainability.
      \end{itemize}
    \item \textbf{Challenges}:
      \begin{itemize}
        \item Data quality, scalability, and cost of adoption.
      \end{itemize}
    \item \textbf{Future Scope}:
      \begin{itemize}
        \item Integration with IoT and real-time systems.
        \item Wider adoption in developing regions.
      \end{itemize}
  \end{itemize}
\end{frame}
