%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large Data Preprocessing with Scikit-Learn}
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large Data Preprocessing for Pima-Indians Diabetis Dataset}

{\tiny (Ref: How To Prepare Your Data For Machine Learning in Python with Scikit-Learn - Jason Brownlee)}
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Read Data}
Import Pima Inidans dataset and split into Features (X) and target (Y)
\begin{lstlisting}
import pandas
import scipy
import numpy

url = "https://raw.githubusercontent.com/jbrownlee/Datasets/master/pima-indians-diabetes.data.csv"
names = ['preg', 'plas', 'pres', 'skin', 'test', 'mass', 'pedi', 'age', 'class']

dataframe = pandas.read_csv(url, names=names)
array = dataframe.values

# separate array into input and output components
X = array[:,0:8]
Y = array[:,8]
\end{lstlisting}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Rescale Data}

	\begin{itemize}
	\item Referred to as normalization and attributes are often rescaled into the range between 0 and 1. 
	\item This is useful for optimization algorithms in used in the core of machine learning algorithms like gradient descent. 
	\item It is also useful for algorithms that weight inputs like regression and neural networks and algorithms that use distance measures like K-Nearest Neighbors.
	\end{itemize}
	
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Rescale Data}
Rescale your data using scikit-learn using the MinMaxScaler class.
\begin{lstlisting}
# Rescale data (between 0 and 1)
from sklearn.preprocessing import MinMaxScaler

scaler = MinMaxScaler(feature_range=(0, 1))
rescaledX = scaler.fit_transform(X)

# summarize transformed data
numpy.set_printoptions(precision=3)
print(rescaledX[0:5,:])
\end{lstlisting}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Rescale Data}
After rescaling you can see that all of the values are in the range between 0 and 1.
\begin{lstlisting}
[[ 0.353  0.744  0.59   0.354  0.     0.501  0.234  0.483]
 [ 0.059  0.427  0.541  0.293  0.     0.396  0.117  0.167]
 [ 0.471  0.92   0.525  0.     0.     0.347  0.254  0.183]
 [ 0.059  0.447  0.541  0.232  0.111  0.419  0.038  0.   ]
 [ 0.     0.688  0.328  0.354  0.199  0.642  0.944  0.2  ]]
\end{lstlisting}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Standardize Data}

	\begin{itemize}
	\item Standardization is a useful technique to transform attributes with a Gaussian distribution and differing means and standard deviations to a standard Gaussian distribution with a mean of 0 and a standard deviation of 1.
	\item It is most suitable for techniques that assume a Gaussian distribution in the input variables and work better with rescaled data, such as linear regression, logistic regression and linear discriminate analysis.
	\end{itemize}

\begin{block}{Intuition}
Rescale (MinMaxScaler) and Standardize (StandardScaler) are easy to mix up: rescaling squeezes every value into a fixed $[0,1]$ box regardless of shape, while standardizing centers the data on 0 and measures everything in ``number of standard deviations'' -- which is why standardized values can still go negative or exceed 1, unlike rescaled ones.
\end{block}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Standardize Data}
Standardize data using scikit-learn with the StandardScaler class.
\begin{lstlisting}
# Standardize data (0 mean, 1 stdev)
from sklearn.preprocessing import StandardScaler

scaler = StandardScaler().fit(X)
rescaledX = scaler.transform(X)

# summarize transformed data
numpy.set_printoptions(precision=3)
print(rescaledX[0:5,:])
\end{lstlisting}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Standardize Data}
The values for each attribute now have a mean value of 0 and a standard deviation of 1.
\begin{lstlisting}
[[ 0.64   0.848  0.15   0.907 -0.693  0.204  0.468  1.426]
 [-0.845 -1.123 -0.161  0.531 -0.693 -0.684 -0.365 -0.191]
 [ 1.234  1.944 -0.264 -1.288 -0.693 -1.103  0.604 -0.106]
 [-0.845 -0.998 -0.161  0.155  0.123 -0.494 -0.921 -1.042]
 [-1.142  0.504 -1.505  0.907  0.766  1.41   5.485 -0.02 ]]
\end{lstlisting}
\end{frame}




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Normalize Data}

	\begin{itemize}
	\item Normalizing in scikit-learn refers to rescaling each observation (row) to have a length of 1 (called a unit norm in linear algebra).
	\item This preprocessing can be useful for sparse datasets (lots of zeros) with attributes of varying scales when using algorithms that weight input values such as neural networks and algorithms that use distance measures such as K-Nearest Neighbors.
	\end{itemize}
	
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Normalize Data}
Normalize data in Python with scikit-learn using the Normalizer class.
 \begin{lstlisting}
# Normalize data (length of 1)
from sklearn.preprocessing import Normalizer

scaler = Normalizer().fit(X)
normalizedX = scaler.transform(X)
# summarize transformed data
numpy.set_printoptions(precision=3)
print(normalizedX[0:5,:])
\end{lstlisting}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Normalize Data}
Each row is rescaled so that the sum of the squares of its values equals 1 -- the row's direction is preserved, only its length changes. A simple worked example, for one row of two features. Applied to every row of \texttt{X}, this is exactly what \texttt{normalizedX} above contains -- each row now has unit length, so no single value in it can exceed 1 in magnitude.
\begin{lstlisting}
row = [4, 3]                     # length = sqrt(4^2 + 3^2) = 5
normalized_row = [4/5, 3/5]      # = [0.8, 0.6]
# check: 0.8^2 + 0.6^2 = 1.0
\end{lstlisting}
\end{frame}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Binarize Data (Make Binary)}

	\begin{itemize}
	\item You can transform your data using a binary threshold. All values above the threshold are marked 1 and all equal to or below are marked as 0.
	\item This is called binarizing your data or threshold your data. It can be useful when you have probabilities that you want to make crisp values. 
	\item It is also useful when feature engineering and you want to add new features that indicate something meaningful.
	\end{itemize}
	
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Binarize Data}
Create new binary attributes in Python using scikit-learn with the Binarizer class.
\begin{lstlisting}
# binarization
from sklearn.preprocessing import Binarizer

binarizer = Binarizer(threshold=0.0).fit(X)
binaryX = binarizer.transform(X)
# summarize transformed data
numpy.set_printoptions(precision=3)
print(binaryX[0:5,:])
\end{lstlisting}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Binarize Data}
You can see that all values equal or less than 0 are marked 0 and all of those above 0 are marked 1.
\begin{lstlisting}
[[ 1.  1.  1.  1.  0.  1.  1.  1.]
 [ 1.  1.  1.  1.  0.  1.  1.  1.]
 [ 1.  1.  1.  0.  0.  1.  1.  1.]
 [ 1.  1.  1.  1.  1.  1.  1.  1.]
 [ 0.  1.  1.  1.  1.  1.  1.  1.]]
\end{lstlisting}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Quick Check: Data Preprocessing}
\begin{block}{Think About It}
You fit a \texttt{StandardScaler} on your training data, then need to preprocess the test set before evaluating your model. Should you call \texttt{scaler.fit(X\_test)} again, or reuse the scaler fitted on the training data?
\end{block}
\pause
\begin{block}{Answer}
Always reuse the scaler fitted on the training data (\texttt{scaler.transform(X\_test)}, no re-fit). Fitting a new scaler on the test set would compute a \emph{different} mean/std from data the model has never seen, so the same raw value could be scaled differently between train and test -- silently corrupting the features the model was trained on. Fitting only ever happens once, on training data; every other split reuses that same fitted scaler.
\end{block}
\end{frame}

