%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}

\begin{center}
{\Large NER with Spacy}
\end{center}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Spacy}
  \begin{itemize}
  \item SpaCy is an open-source python library for NLP written in Python and Cython. 
	\item Offers pre-trained models for multi-language NER, as well as allowing developers to train and deploy custom NER models on domain specific corpuses. 
	\item SpaCy models are designed to be production-ready.
	\item Uses 1D residual convolutional neural networks (CNN) and incremental parsing with Bloom embeddings for NER
  \end{itemize}
	
	{\tiny (Ref: Google Cloud AI Hub Named Entity Recognition using Spacy and Tensorflow)}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Spacy Default NER}
Identifies:
  \begin{itemize}
  \item PERSON:	People, including fictional
	\item ORG:	Companies, agencies, institutions, etc
	\item GPE:	Countries, cities, states
	\item PRODUCT:	Objects, vehicles, foods, etc. (Not services.)
	\item DATE:	Absolute or relative dates or periods
	\item TIME:	Times smaller than a day
	\item PERCENT:	Percentage, including ”%“
	\item MONEY:	Monetary values, including unit
	\item QUANTITY:	Measurements, as of weight or distance
	\item \ldots
  \end{itemize}
	
	{\tiny (Ref: Google Cloud AI Hub Named Entity Recognition using Spacy and Tensorflow)}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Spacy Default NER}

\begin{lstlisting}
import spacy
nlp = spacy.load('en_core_web_sm')

doc = nlp("Indians spent over $71 billion on clothes in 2018")
 
for ent in doc.ents:
    print(ent.text, ent.label_)
		
Indians NORP
over $71 billion MONEY
2018 DATE

spacy.explain("NORP")
Output: `Nationalities religious or political groups'
\end{lstlisting}
	
	{\tiny (Ref: spaCy Tutorial to Learn and Master Natural Language Processing (NLP) - Prateek Joshi - Analytics Vidhya)}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}

\begin{center}
{\Large Custom NER}
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Training your own NER}
  \begin{itemize}
  \item Either rule-based or machine learning (ML) based.
	\item Requires a large amount of labeled training data
	\item ML approaches from scratch: Hidden Markov Models, Maximum Entropy, and Conditional Random Fields, as well as deep learning approaches with Recurrent Neural Networks, such as Seq2Seq
	\item Need sentence inputs and annotated sentence outputs. 
	\item May also involve additional feature engineering
	\item Some libraries like Spacy and Stanford allow training of custom NER.
  \end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{NER Datasets}
  \begin{itemize}
  \item Domain-specific (i.e. Twitter, biomedical, advertising, news).
	\item i2b2 - Medication, treatments, diseases, risk factors, and medications
	\item CoNLL 2003 - English and german news articles annotated with location, organization, person, and miscellaneous
  \end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{NER Evaluation metrics}
  \begin{itemize}
  \item NER is most commonly evaluated with precision, recall, and F1-score.
  \end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Training your own system}
  \begin{itemize}
  \item The feature extraction works almost identical as the one implemented in the Training a Part-Of-Speech Tagger, except need to add many features.
  \item Since the previous IOB tag is a very good indicator of what the current IOB tag is going to be, we have included the previous IOB tag as a feature.
  \item Spacy Provides Gold Parse method
  \item CRF++ can be used to generate custom NER tags.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}

\begin{center}
{\Large Spacy Custom NER -TBD-}
\end{center}
\end{frame}


