%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large Python for AI}
\end{center}

% \tiny{ (Ref: https://automatetheboringstuff.com/ )}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{What's going on in AI?}
\begin{itemize}
\item Companies/platforms
\item Achievements, 
\item and products to use. 
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{What Languages Are Used for Building AI?}
\begin{itemize}
\item LISP (used to be?) one of the most popular languages for creating AI. Dynamic typing and lists.
\item C/C++ is used for building performance AI, but not so easy to program.
\item  Java is not as fast as C but its portability and built-in types make Java a choice of many developers. 
\item And finally, there is Python. 
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Why Python scores?}
\begin{itemize}
\item Simple, clean syntax
\item Many powerful libraries
\item Many domains
\item Language interoperability
\item Free!!
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Libraries}
Related to AI-ML-DL
\begin{itemize}
\item Numpy, Scipy for mathematics
\item Pandas from data processing
\item Matplotlib for visualizations
\item Scikit Learn for Machine Learning
\item and many more
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Machine Learning Topics with Python}
\begin{itemize}
\item Regression
\item Classification
\item Clustering
\item Dimension Reduction
\item and many more
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Deep Learning with Python}
\begin{itemize}
\item Neural Networks
\item Convolution Neural Network
\item Recurrent Neural Network
\item Libraries: Tensorflow, Keras, Pytorch, etc
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{What Next?}
Resources:
\begin{itemize}
\item Google Developers Python Course (highly recommended for visual learners)
\item An Introduction to Python for Scientific Computing (from UCSB Engineering), by M. Scott Shell (a great scientific Python intro ~60 pages)
\item Datacamp online course
\item Learn X in Y Minutes (X = Python) 
\item Book ``Python The Hard Way'' by Zed A. Shaw (a 30 minute crash course)
\end{itemize}
If you think, you need a human to teach, then join the workshop !!!
\end{frame}