%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large Assignments}
\end{center}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Find similar company names}
Quite often, we have strings that refer to the same entity but have different string representations (e.g., McDonald's vs McDonalds vs McDonald). Need to find such similar entries in our data.
  \begin{itemize}
  \item Step 1: Read the data into a list of strings. The list of unique restaurant names from the NYC Restaurant Inspection dataset (data/restaurant-names.txt) and a list of bank names (data/bank-names.txt)
  \item Step 2: How to compare two strings and find their similarity.
  \item Step 3: Write a function that takes as input a company name, and returns back a list of matching company names, together with their similarity.
  \item Step 4: In the final step, we will just perform the similarity computation across all companies in the dataset.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{STEP 1}
Read the list of names from a file and create a list of names
  \begin{lstlisting}
# STEP 1: Read the list of names 
path = 'data/'
filename = path + 'restaurant-names.txt'
  \end{lstlisting}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{STEP 2}
Computing the similarity between two strings. There are many ways that we can calculate the similarity between two strings. 
  \begin{itemize}
  \item We will use the q-gram similarity metric. 
  \item What is a q-gram? Simply a sequence of q-consecutive characters in the string. 
  \item For example, the name "Panos" has the following 2-grams: "Pa", "an", "no", "os". 
  \item Strings that share a large number of q-grams are often similar. Compute the Jaccard coefficient between these sets. 
  \end{itemize}
  \begin{lstlisting}
# STEP 2: Now we implement our similarity function

# This returns a list of q-grams for a name
# getQgrams("Panos", 1) should return ["P", "a", "n", "s"]
# getQgrams("Panos", 2) should return ["Pa", "an"]
def getQgrams(name, q):
    pass
  \end{lstlisting}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{STEP 3}
  \begin{lstlisting}
# STEP 3: We now create a function that accepts a company name
# and a list of companies, and computes their similarity
# We have a threshold parameter (by default set to be 0.7)
# that restricts the results to only string pairs with similarity
# above the threshold

# This function takes as input two names, computes their qgrams, and then computes
# the Jaccard coefficient (=intersection/union) of the two sets of qgrams
def computeSimilarity(name1, name2, q):
    # your code here
    pass
    
print computeSimilarity("Peter", "Pete", 2)
  \end{lstlisting}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{STEP 4}
Perform the similarity computation across all companies in the dataset.
  \begin{lstlisting}
# STEP 4: We are almost done. We now just go through all the companies in the list
# and we call the companySimilarity function that computes the similar company names
# for all the companies in the list. We store the results in a dictionary, with the 
# key being the company name, and the value being a "list of tuples" with the 
# similar company names and the corresponding similarity value.

  \end{lstlisting}
\end{frame}
