%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large What's next?}
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}\frametitle{Recap}

\begin{itemize}
\item A scripting language, Python, is suitable:
	\begin{itemize}
	\item for embedding
	\item for writing small unstructured scripts,
	\item for ``quick and dirty'' programs.
	\end{itemize}
\item Not JUST ANY a scripting language:
	\begin{itemize}
	\item Python scales.
	\item Python encourages us to write code that is clear and well-structured.
	\end{itemize}
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}\frametitle{Recap}

\begin{itemize}
%\item Interpreted. 
%\item Modules are automatically compiled (to .pyc) when imported, but may also be explicitly compiled.
\item Provides an interactive command line and interpreter shell.
\item Dynamic:
	\begin{itemize}
	\item Types are bound to values, not to variables.
	\item Values are inspect-able.
	\item You can list the methods supported by any given object.
	\end{itemize}
\item Strongly typed at run-time, not compile-time. 
\item Objects (values) have a type, but variables do not (that's why you can assign to any value type).
\item Reasonably high level -- High level built-in data types; 
\item High level control structures (for walking lists and iterators, for example).
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}\frametitle{Recap}

\begin{itemize}
\item Object-oriented -- Almost everything is an object.
\item Readability, locate-ability, modifiability.
\item Indented block structure -- ``Python is pseudo-code that runs.''
\item Embedding and extending Python:
\begin{itemize}
\item  Embed the Python interpreter in C/C++ applications (FreeCAD)
\item  Modules and objects implemented in C/C++ (SWIG)
\item Cython enables us to generate C code from Python
\end{itemize}
\item Automatic garbage collection
\end{itemize}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}\frametitle{The Zen of Python (by Tim Peters)}
\scriptsize
\begin{itemize}
\item Beautiful is better than ugly.
\item Explicit is better than implicit.
\item Simple is better than complex.
\item Complex is better than complicated.
\item Flat is better than nested.
\item Sparse is better than dense.
\item Readability counts.
\item Special cases aren't special enough to break the rules.
\item Practicality beats purity.
\item Errors should never pass silently, unless explicitly silenced.
\item In the face of ambiguity, refuse the temptation to guess.
\item There should be one-- and preferably only one --obvious way to do it, although that way may not be obvious at first unless you're Dutch.
\item Now is better than never, although never is often better than *right* now.
\item If the implementation is hard to explain, it's a bad idea.
\item If the implementation is easy to explain, it may be a good idea.
\item Namespaces are one honking great idea -- let's do more of those!
\end{itemize}
(Presented at a talk on the History of Python by Guido van Rossum in 2005 at EuroPython in Gothenburg, Sweden)
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}\frametitle{Bye}
{\large Life is Short}
(You Need Python)
- Bruce Eckel (Thinking in C++)
\end{frame}


% Ideas?

