%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large Vertex AI}
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Introduction}

\begin{itemize}
\item Vertex AI is a unified machine learning (ML) platform that helps you build, deploy, and scale ML models.
\item It provides a single, integrated experience for the entire ML lifecycle, from data preparation to model training to deployment.
\item Vertex AI is built on top of Google Cloud's infrastructure, so you can be confident that your models will be scalable, reliable, and secure.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Concepts}

\begin{itemize}
\item Datasets: Datasets are the foundation of Vertex AI. They store your data in a format that Vertex AI can understand and use.
\item Models: Models are the output of the ML training process. They are mathematical representations of the relationships between your data and the desired output.
\item Endpoints: Endpoints are the way that you make predictions from your models. They are hosted in the cloud and can be accessed by your applications.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Framework}

\begin{itemize}
\item AutoML: AutoML is a codeless service that can automatically train and deploy ML models for you.
\item Kubeflow Pipelines: Kubeflow Pipelines is a framework for building and deploying ML pipelines.
\item TensorFlow: TensorFlow is an open-source framework for machine learning.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Applications}

\begin{itemize}
\item Image classification: Vertex AI can be used to classify images, such as identifying objects in a photo.
\item Natural language processing: Vertex AI can be used to process natural language, such as understanding the meaning of a sentence.
\item Fraud detection: Vertex AI can be used to detect fraud, such as identifying fraudulent credit card transactions.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{Conclusion}

\begin{itemize}
\item Vertex AI is a powerful ML platform that can help you build, deploy, and scale ML models.
\item It is a comprehensive platform that provides a single, integrated experience for the entire ML lifecycle.
\item Vertex AI is built on top of Google Cloud's infrastructure, so you can be confident that your models will be scalable, reliable, and secure.
\end{itemize}

\end{frame}
