%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]\frametitle{}
\begin{center}
{\Large Introduction to Data Structures and Algorithms}
\end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
	\frametitle{What are Data Structures?}
		\begin{itemize}
			\item Different ways in which data is organized
			\item So as to be effectively used in Computer Algorithms
			\item E.g. its easier to find something if it is ordered or organized in buckets with similar properties.
		\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
	\frametitle{What is an Algorithm?}
		\begin{itemize}
			\item A set of steps to accomplish a task
			\item E.g. Making a dish involves, several sequential/parallel steps.
			\item Good algorithms aim for Correctness and Efficiency
		\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
	\frametitle{Why study?}
		\begin{itemize}
			\item Wide variety of applications
			\item Need optimized, fast, correct, robust solutions.
			\item Trade-off of Space and Time
			\item Skill : problem solving
			\item Programming language agnostic.
		\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
	\frametitle{Types of Data Structures}
		\begin{itemize}
			\item Primitive/predefined: integer, float, character, string, boolean, etc.
			\item Custom/user-defined: 
					\begin{itemize}
					\item Linear
					\begin{itemize}
						\item Static: Array
						\item Dynamic: Linked lists, Stacks, Queues
					\end{itemize}					
					\item Non-Linear: Trees, Graphs
				\end{itemize}
		\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
	\frametitle{Types of Algorithms}
		\begin{itemize}
			\item Recursive: function calls itself
			\item Divide and Conquer: Divide into sub-problems, solve them and compose solutions
			\item Dynamic Programming: Memorize solutions to subproblem to be reused and not recompute
			\item Greedy: By choosing local optimum solution, hopes to find global optimum solution.
			\item Randomized: Different random paths taken
			\item Brute Force: Try all possibilities until solution is found.
		\end{itemize}
\end{frame}