\section{Intermezzo}
\label{sec:intermezzo}

\begin{frame}
	\frametitle{Intermezzo: MapReduce}
	
	\begin{center}
	\input{figures/tikz/mapreduce.tex}
	\hspace*{15pt}\hbox{\scriptsize Image Based on original image by :\thinspace{\itshape Clém IAGL}}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{What is MapReduce?}
	\begin{itemize}
		\item Original paper by Google: ``MapReduce: Simplified Data Processing on Large Clusters''.
		\item A technique they used to process their massive amounts of data streaming in.
			\pause
		\item The main idea?
			\begin{itemize}
				\item Split up the data.
				\item Apply some basic function to it (Map).
				\item Merge the results (Reduce).
			\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Example}
		\begin{problemblock}{A word count}
			Imagine I have a several books and what to know how often certain words appear in the books.	\\
		\end{problemblock}	
		\pause
		\begin{answerblock}{MapReduce}
			The MapReduce strategy is to:
		\begin{enumerate}
			\item Split the books into pages.
			\item Pass every page to a different machine and ask it to count the words, returning a \texttt{dict} from a word
				to a count.
			\item Aggregate these results.
		\end{enumerate}
			
		\end{answerblock}
		\pause
			\begin{exampleblock}{Why is this nice?}
				We can easily \textit{parallelise} our problem over many machines (they do the heavy-lifting) and then simply
				aggregate the intermediate results (relatively quick/cheap operation).	
			\end{exampleblock}	
\end{frame}

