\begin{frame}
	\frametitle{BogoSort}
		\begin{block}{BogoSort}
			The easiest form of sorting: Just try all permutations of the list.
		\end{block}	
		\pause
		\begin{questionblock}{Time complexity?}
			\begin{enumerate}[A.]
				\item $\Theta(n)$
				\item $\Theta(n^2)$
				\item $\Theta(2^n)$
				\item $\Theta(n^n)$
				\item $\Theta(n!)$
				\item I don't know
			\end{enumerate}
		\end{questionblock}
	
\end{frame}

\begin{frame}
	\frametitle{Some questions and answers}
\begin{block}{The scope of our problem}
	Some observations and questions.
	\begin{itemize}
		\item The solution space is huge: $n!$ for $n$ items.
			\pause
		\item The use case is not uncommon, so we hope/look for an efficient (polynomial time) algorithm.
			\pause
		\item Humans have a lot of experience sorting items, can we use those strategies?
			\pause
		\item Or are there methods that computers excel at, that would be terrible for humans?
	\end{itemize}
\end{block}	
\pause
\begin{answerblock}{Yes!}
	\begin{itemize}
		\item There are polynomial time algorithms,
			\pause
		\item We can translate our own methods,
		\item But computers can definitely do better.
	\end{itemize}
\end{answerblock}
\end{frame}

\begin{frame}
	\frametitle{Use case?}
	\begin{questionblock}{Why?}
		What do we use sorting for?
	\end{questionblock}
	\pause
	\begin{answerblock}{Many things!}
		\begin{itemize}
			\item Books in the library.
			\item Exams for an exam review.
			\item Closest pair of points.
			\item Convex hulls.
			\item Efficient look-ups in lists that do not change.
			\item And many many more (see also later lectures in graph algorithms).
		\end{itemize}	
	\end{answerblock}
\end{frame}
