\subsection{Week 8}
\label{sub:week_8}

\subsubsection{Lecture 13: P vs NP}
\label{sub:lecture_13}

\hfill\\
\textbf{Learning Objectives:}\\
This lecture partially concerns \cref{lo:graph} of the course, but most of this lecture is not exam-material this year.
More specifically after this lecture the student is able to:
\begin{itemize}
	\item describe the class of problems called P.
	\item describe the class of problems called NP.
	\item describe the relation between P and NP.
	\item show that certain problems are part of P and/or NP.
\end{itemize}

\hfill\\
\textbf{Intermezzo:}\\
Approximation algorithms.

\hfill\\
\textbf{Lecture problem:}\\
Hamiltonian cycle, TSP.

\hfill\\
\textbf{Lecture plan:}\\
\begin{itemize}
	\item[5 min] Notes on the admin related side \& questions from students.
	\item[5 min] Recap of last lecture, graphs.
	\item[5 min] Introduce problem of ham cycle.
	\item[5 min] Show how we can verify a solution.
	\item[10 min] Introduce classes P and NP.
	\item[10 min] Go over some previous problems and classify them.
	\item[5 min] Explain ham cycle is in NP, but we do not know if it is in P.
	\item Break time
	\item[5 min] Intermezzo
	\item[10 min] Introduce TSP (open day like).
	\item[20 min] Solving TSP with additional restrictions.
	\item[5 min] End with spiel about how P vs NP is very cool and we need mathematicians to help us solve it :)
	\item[5 min] Short survey about the lecture: what was unclear, what was good, what could be better?
\end{itemize}

\newpage
\subsubsection{Lecture 14: Exam prep}
\label{sub:lecture_14}

\hfill\\
\textbf{Learning Objectives:}\\
?
\hfill\\
\textbf{Intermezzo:}\\
None

\hfill\\
\textbf{Lecture problem:}\\
None

\hfill\\
\textbf{Lecture plan:}\\
\begin{itemize}
	\item[5 min] Notes on the admin related side \& questions from students.
	\item[80 min] Whatever the students want.
	\item[5 min] Short survey about the lecture: what was unclear, what was good, what could be better?
\end{itemize}

\newpage
\subsubsection{Tutorial 7: Exam prep}
\label{ssub:tutorial_7}

\hfill\\
\textbf{Plan:}\\
\begin{itemize}
	\item[85 min] Whatever the students want.
	\item[5 min] Short survey about the tutorial: what was unclear, what was good, what could be better?
\end{itemize}


\newpage
\subsubsection{Assignment 7: Not applicable}
\label{ssub:assignment_7}

Does not exist. \unsure{ We cannot get the students feedback in time, as the exam is on the Monday of week 9. But
perhaps we can think of something nice to put her anyway?}
