\subsection{Week 3}
\label{sub:week_3}

\subsubsection{Lecture 5: Stacks and queues}
\label{sub:lecture_5}

\hfill\\
\textbf{Learning Objectives:}\\
This lecture concerns \cref{lo:stack,lo:queue} of the course, more specifically after this lecture the student is able
to:
\begin{itemize}
	\item define the interface of a stack and of a queue.
	\item describe the differences between a stack and a queue.
	\item implement one interface in terms of the other (i.e. a stack from 2 queues or a queue from 2 stacks).
	\item select the right one (either a stack or a queue) given a use case.
\end{itemize}

\hfill\\
\textbf{Intermezzo:}\\
Routers: fifo in routing.

\hfill\\
\textbf{Lecture problem:}\\
Parsing LaTeX (stacks) and the supermarket registry (queues).

\hfill\\
\textbf{Lecture plan:}\\
\begin{itemize}
	\item[5 min] Notes on the admin related side \& questions from students.
	\item[5 min] Recap of last lecture, reminder of different types of lists.
	\item[5 min] Introduce problem of parsing TeX.
	\item[5 min] Stack properties.
	\item[5 min] Stack implementation: linked-list based.
	\item[10 min] Using a stack to parse our TeX.
	\item[5 min] Introduce problem of supermarket queuing (and maybe traffic jams).
	\item[5 min] Queue properties.
	\item Break time
	\item[5 min] Intermezzo
	\item[5 min] Queue implementation: linked-list based.
	\item[5 min] Solving the problem of the supermarket.
	\item[10 min] Queue implementation: circularly-linked list.
	\item[5 min] Whoops, we made a Deque now.
	\item[10 min] Using a Stack as a Queue.
	\item[5 min] Short survey about the lecture: what was unclear, what was good, what could be better?
\end{itemize}

\newpage
\subsubsection{Lecture 6: Sorting}
\label{sub:lecture_6}

\hfill\\
\textbf{Learning Objectives:}\\

This lecture concerns \cref{lo:sort} of the course, more specifically after this lecture the student is able
to:
\begin{itemize}
	\item compare and implement bubble sort, selection sort, merge sort, bucket sort, and quicksort.
	\item combine different sorting strategies.
	\item select the right type of sorting given a use case.
\end{itemize}

\hfill\\
\textbf{Intermezzo:}\\
Algorithms to Live by.

\hfill\\
\textbf{Lecture problem:}\\
Remember Binary Search? Only good if the list is sorted. \improvement{Needs a nice story}

\hfill\\
\textbf{Lecture plan:}\\
\begin{itemize}
	\item[5 min] Notes on the admin related side \& questions from students.
	\item[5 min] Recap of binary search.
	\item[5 min] Introduction to today's topic: sorting.
	\item[10 min] Bubble-sort: idea, algorithm, and analysis.
	\item[15 min] Selection-sort: idea, algorithm, and analysis.
	\item[5 min] Bucket-sort: idea, algorithm, and analysis.
	\item Break time
	\item[5 min] Intermezzo
	\item[15 min] Merge-sort: idea, algorithm, and analysis (master method).
	\item[15 min] Quick-sort: idea, algorithm, and analysis.
	\item[5 min] Recap: what did we see, how are they different?
	\item[5 min] Short survey about the lecture: what was unclear, what was good, what could be better?
\end{itemize}

\newpage
\subsubsection{Tutorial 3: Stacks, Queues, Sorting}
\label{ssub:tutorial_3}

\hfill\\
\textbf{Plan:}\\
\begin{itemize}
	\item[15 min] Exercise C6.20 from the book.
	\item[10 min] Exercise C6.26 from the book.
	\item[15 min] Implement Selection-sort.
	\item[5 min] Sorting with a key-function in python.
	\item Break time
	\item[5 min] Questions about popping/pushing from/to stacks/queues.
	\item[5 min] Analyse C6.26.
	\item[10 min] Exercise R12.10.
	\item[5 min] Exercise R12.15.
	\item[15 min] Analysing random quick-sort? \unsure{I might also be interested in teaching something like 12.4.1}
	\item[5 min] Short survey about the tutorial: what was unclear, what was good, what could be better?
\end{itemize}


\newpage
\subsubsection{Assignment 3: Deadline Sunday before week 4}
\label{ssub:assignment_3}

\hfill\\
\textbf{Analysis:}\\
\begin{description}
	\item[] Give T(n) and determine big-Oh. \unsure{Look into ways to auto-grade without making it MC.}
		\begin{description}
			\item[10 min] 2 simple recursions.
			\item[10 min] More complicated recursion.
			\item[15 min] Something with stacks.
			\item[15 min] Something with deques.
			\item[10 min] Sorting algorithm A.
			\item[10 min] Sorting algorithm B.
			\item[10 min] Sorting algorithm C.
		\end{description}
	\item[15 min] 5 MC (exam-level) questions.
	\item[10 min] 2 rec. equations to solve using the master method.
	\item[10 min] 1 rec. equations to solve using repeated unfolding + induction.
	\item[10 min] Big-Theta proof.
\end{description}

\hfill\\
\textbf{Implementation:}\\
\begin{description}
	\item[30 min] Implement a Stack with Queues (explained in lecture).
	\item[45 min] Implement a Queue with Stacks (not explained in lecture).
	\item[45 min] Implement merge-sort.
	\item[45 min] Implement quick-sort.
	\item[30 min] \hfill\\\unsure{Some problem that needs solving using only queues and stacks, students should select
		which.}
	\item[45 min] \hfill\\\unsure{A sorting algorithm not explained in the lecture.}
	\item[30 min] Implement sorting on a linked-list structure of some kind (partially recap).
\end{description}
