\href{\webworkurl}{Webwork}


\begin{enumerate}
\item Let $M=\begin{pmatrix}
2 & 1 \\
0 & 2
\end{pmatrix}$.  Find all eigenvalues of $M$.  Does $M$ have two independent\footnote{Independence of vectors is  explained \hyperref[independent]{here}.} eigenvectors?  Can $M$ be diagonalized?

\phantomnewpage

\item Consider $L \colon \Re^2\rightarrow \Re^2$ with $L\colvec{x\\y}
=\colvec{x\cos \theta +y\sin \theta\\ -x\sin \theta + y\cos \theta}$.
\begin{enumerate}
\item Write the matrix of $L$ in the basis $\colvec{1\\0}, \colvec{0\\1}$.

\item When $\theta\neq 0$, explain how $L$ acts on the plane.  Draw a picture.

\item Do you expect $L$ to have invariant directions?

\item Try to find real eigenvalues for $L$ by solving the equation 
\[
L(v)=\lambda v.
\]

\item Are there complex eigenvalues for $L$, assuming that $i=\sqrt{-1}$ exists?
\end{enumerate}

\phantomnewpage

\item Let $L$ be the linear transformation $L \colon \Re^3\rightarrow \Re^3$ given by 
$$L\colvec{x\\y\\z}=\colvec{ x+y \\  x+z \\ y+z}.$$  
Let $e_i$ be the vector with a one in the $i$th position and zeros in all other positions.
\begin{enumerate}
\item Find $Le_i$ for each $i$.
\item Given a matrix $M=\begin{pmatrix}
m^1_1 & m^1_2 & m^1_3\\
m^2_1 & m^2_2 & m^2_3\\
m^3_1 & m^3_2 & m^3_3\\
\end{pmatrix}$, what can you say about $Me_i$ for each $i$?
\item Find a $3\times 3$ matrix $M$ representing $L$.  Choose three nonzero vectors pointing in different directions and show that $Mv=Lv$ for each of your choices.
\item Find the eigenvectors and eigenvalues of \(M.\)

\end{enumerate}

\phantomnewpage

\item Let $A$ be a matrix with eigenvector $v$ with eigenvalue $\lambda$. Show that $v$ is also an eigenvector for $A^2$ and what is its eigenvalue? How about for $A^n$ where $n \in \mathbb{N}$? Suppose that $A$ is invertible. Show that $v$ is also an eigenvector for $A^{-1}$.

\item A \emph{projection}\index{Projection} is a linear operator $P$ such that $P^2 = P$. Let $v$ be an eigenvector with eigenvalue $\lambda$ for a projection $P$, what are all possible values of $\lambda$? Show that every projection $P$ has at least one eigenvector.

Note that every complex matrix has at least 1 eigenvector, but you need to prove the above for \emph{any} field.

\end{enumerate}

\phantomnewpage