
\subsection*{Hint for Review Problem~\ref{injectivekernalprob} }

%%%Insert this to get the typewriter font so it looks like a real movie script
{\ttfamily
\fontdimen2\font=0.4em
\fontdimen3\font=0.2em
\fontdimen4\font=0.1em
\fontdimen7\font=0.1em
\hyphenchar\font=`\-


\hypertarget{scripts_kernel_range_nullity_rank_hint}{Lets work through this} problem. 

Let $L \colon V\rightarrow W$ be a linear transformation.  Show that $\ker L=\{0_V\}$ if and only if $L$ is one-to-one:
\begin{enumerate}
\item First, suppose that \(\ker L=\{0_V\}\). Show that \(L\) is one-to-one. 

Remember what one-one means, it means whenever $L(x) = L(y)$ we can be certain that $x=y$. While this might seem like a weird thing to require this statement really means that each vector in the range gets mapped to a unique vector in the range. 

We know we have the one-one property, but we also don't want to forget some of the more basic properties of linear transformations namely that they are linear, which means $L(ax+by) = aL(x) + bL(y)$ for scalars $a$ and $b$. 

What if we rephrase the one-one property to say whenever $L(x) -L(y) = 0$ implies that $x-y = 0$? Can we connect that to the statement that $\ker L=\{0_V\}$? Remember that if $L(v) = 0 $ then $v \in \ker L=\{0_V\}$.

\item Now, suppose that \(L\) is one-to-one. Show that \(\ker L=\{0_V\}\). That is, show that \(0_V\) is in \(\ker L\), and then show that there are no other vectors in \(\ker L\).

What would happen if we had a nonzero kernel? If we had some vector $v$ with $L(v) = 0 $ and $v \not = 0 $, we could try to show that this would contradict the given that L is one-one. If we found $x$ and $y$ with $L(x) = L(y)$, then we know $x=y$. But if $L(v) = 0 $ then $L(x) + L(v) = L(y)$. Does this cause a problem?

\end{enumerate}




} % Closing braket for font

%\newpage
