
\subsection*{Invertibility Conditions}

%%%Insert this to get the typewriter font so it looks like a real movie script
{\ttfamily
\fontdimen2\font=0.4em
\fontdimen3\font=0.2em
\fontdimen4\font=0.1em
\fontdimen7\font=0.1em
\hyphenchar\font=`\-


\hypertarget{scripts_kernel_range_nullity_rank_inv_cond}{Here I am going to discuss some of} the conditions on the invertibility of a matrix stated in \hyperref[theorem:invertibility]{Theorem~\ref*{theorem:invertibility}}. Condition~1 states that $X = M^{-1}V$ uniquely, which is clearly equivalent to~4. Similarly, every square matrix $M$ uniquely corresponds to a linear transformation $L \colon \mathbb{R}^n \rightarrow \mathbb{R}^n$, so condition~3 is equivalent to condition~1.

Condition~6 implies~4 by the adjoint construct the inverse, but the converse is not so obvious. For the converse (4 implying~6), we refer back the proofs in Chapter~18 and~19. Note that if $\det M = 0$, there exists an eigenvalue of $M$ equal to~$0$, which implies $M$ is not invertible. Thus condition~8 is equivalent to conditions~4,~5,~9, and~10.

The map $M$ is injective if it does not have a null space by definition, however eigenvectors with eigenvalue~$0$ form a basis for the null space. Hence conditions~8 and~14 are equivalent, and~14,~15, and~16 are equivalent by the \hyperref[dimension_formula]{Dimension Formula} (also known as the Rank-Nullity Theorem).

Now conditions~11,~12, and~13 are all equivalent by the definition of a basis. Finally if a matrix $M$ is not row-equivalent to the identity matrix, then $\det M = 0$, so conditions~2 and~8 are equivalent.

} % Closing braket for font

%\newpage
