\input{../shared/shared}

\renewcommand{\course}{Robotics}
\renewcommand{\coursepicture}{roboticsLecture}
\renewcommand{\coursedate}{Winter 2014}
\renewcommand{\exnum}{11}

\exercises

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\exsection{Local linearization and Algebraic Riccati equation}

Slide 08:02 describes the cart pole, which is similar to the
Segway-type system of Exercise 6, but a little simpler. We'll solve
the cart pole in this exercise.

The state of the cart-pole is given by $x = (p,\dot p,
\theta,\dot\theta)$, with $p\in\RRR$ the position of the cart,
$\theta\in\RRR$ the pendulums angular deviation from the upright
position and $\dot p,\dot \theta$ their respective temporal
derivatives. The only control signal $u\in\RRR$ is the force applied on the
cart. The analytic model of the cart pole is
\begin{align}
\ddot\theta
 &= \frac{g\sin(\theta) + \cos(\theta)
         \left[-c_1 u - c_2\dot{\theta}^2\sin(\theta)\right]}
    {\frac{4}{3}l - c_2\cos^2(\theta)}\\
\ddot p
 &= c_1 u
  + c_2 \left[\dot\theta^2\sin(\theta) - \ddot\theta\cos(\theta)\right]
\end{align}
with $g = 9.8ms^2$ the gravitational constant, $l = 1m$ the pendulum
length and constants $c_1 = (M_p + M_c)^{-1}$ and $c_2 = l M_p(M_p +
M_c)^{-1}$ where $M_p = M_c = 1kg$ are the pendulum and cart masses
respectively.


a) Derive the local linearization of these dynamics around
$x^*=(0,0,0,0)$. The eventual dynamics should be in the form
\begin{align*}
\dot x = A x + B u
\end{align*}
Note that
\begin{align*}
A = \mat{cccc}{0 & 1 & 0 & 0 \\
 \frac{\del\ddot p}{\del p} & \frac{\del\ddot p}{\del\dot p} &
 \frac{\del\ddot p}{\del\t} & \frac{\del\ddot p}{\del\dot\t} \\
 0 & 0 & 0 & 1 \\
 \frac{\del\ddot \t}{\del p} & \frac{\del\ddot \t}{\del\dot p} &
 \frac{\del\ddot \t}{\del\t} & \frac{\del\ddot \t}{\del\dot\t} }
\comma
B = \mat{c}{0 \\
\frac{\del\ddot p}{\del u} \\
0 \\
\frac{\del\ddot\t}{\del u}}
\end{align*}
where all partial derivatives are taken at the point $p=\dot
p=\t=\dot\t=0$.

The solution (to continue with the other parts) is
\begin{align*}
A = \mat{cccc}{0 & 1 & 0 & 0 \\
  0 & 0 & -\frac{g}{\frac{4}{3} l - c_2} & 0 \\
 0 & 0 & 0 & 1 \\
 0 & 0 & \frac{g}{\frac{4}{3} l - c_2} & 0 }
\comma
B = \mat{c}{0 \\
c_1+\frac{c_1 c_2}{\frac{4}{3} l - c_2} \\
0 \\
\frac{-c_1}{\frac{4}{3} l - c_2}}
\end{align*}


b) We assume a stationary infinite-horizon cost function of the form
\begin{align*}
J^\pi
&= \int_0^\infty c(x(t),u(t))~ dt \\
c(x,u)
&= x^\T Q x + u^\T R u \\
& Q = \diag(c, 0, 1, 0) \comma R = \Id ~.
\end{align*}
That is, we penalize position offset $c\norm{p}^2$ and pole angle
offset $\norm{\t}^2$. Choose $c=\r=1$ to start with.

Solve the Algebraic Riccati equation
\begin{align*}
0 = A^\T P + P^\T A - P B R^\1 B^\T P + Q
\end{align*}
by initializing $P=Q$ and iterating using the following iteration:
\begin{align*}
P_{k\po} = P_k + \e [A^\T P_k + P_k^\T A - P_k B R^\1 B^\T P_k + Q]
\end{align*}
Choose $\e=1/1000$ and iterate until convergence. Output the gains $K=
-R^\1 B^\T P$. (Why should this iteration converge to the solution of
the ARE?)

c) Solve the same Algebraic Riccati equation by calling the @are@
routine of the octave control package (or a similar method in Matlab).
For Octave, install the Ubuntu packages @octave3.2@, @octave-control@,
and @qtoctave@, perhaps use @pkg load control@ and @help are@ in
octave to ensure everything is installed, use %
@P=are(A,B*inverse(R)*B',Q)@ to solve the ARE. Output $K= -R^\1 B^\T P$
and compare to b).

(I found the solution $K=(1.0000, 3.8271, 48.3070,17.615)$.)

d) Implement the optimal Linear Quadratic Regulator $u^*= -R^\1 B^\T P
x$ on the cart pole simulator. In the Robotics code base, path
@07-cartPole@ there is some code -- but the main.problem.cpp was
outdated. Please use the one from the website or email.

Simulate the optimal LQR and test it for various noise levels.



\exerfoot



