\begin{frame}[fragile]
  \frametitle{Variadic functions, I}
  Functions like \texttt{max} and \texttt{min} take a variable number of arguments.

  \+
  That is possible with user-defined functions too.

  \+
  If the last argument is prefixed with a \texttt{*} character,
  Python will make that argument into a \emph{tuple} of arguments
  passed to the function. (Except for the ones that have already been
  assigned names.)
\end{frame}


\begin{frame}[fragile,fragile]
  \frametitle{Variadic functions, II}
\begin{lstlisting}
>>> def varfn1(*args):
...   print(args)
>>> varfn1(1,2,3)
(1, 2, 3)
\end{lstlisting}

\begin{lstlisting}
>>> def varfn2(a, b, *rest):
...   print(rest)
>>> varfn1(1,2,3)
(3,)
\end{lstlisting}
\end{frame}


\begin{frame}[fragile]
  \frametitle{Variadic functions, III}
  You can call a function with an argument list that is only
  determined at run time.

  \+
  The unary \texttt{*} operator takes any sequence and makes it
  into a function argument list:
\begin{lstlisting}
>>> L = [1, 2, 3]
>>> max(*L)
3
\end{lstlisting}
\end{frame}


\begin{frame}[fragile]
  \frametitle{Keyword arguments, I}
  Python lets you catch arbitrary named arguments.

  \+
  Prefix the very last argument name with \texttt{**}, and Python
  will make it into a dictionary: keys are actual argument names and
  dictionary values are actual argument values.

  \+
\begin{lstlisting}
>>> def kwfn1(**kwargs):
...   print(kwargs)
>>> kwfn1(a=1, b=2)
{'a':1, 'b':2}
\end{lstlisting}
\end{frame}


\begin{frame}[fragile,fragile]
  \frametitle{Keyword arguments, II}
  Keyword argument can be combined with positional arguments:
\begin{lstlisting}
>>> def kwfn1(x, y, **kwargs):
...   print("x=%s y=%s kwargs=%s" % (x, y, kwargs))
>>> kwfn1(0, 42, a=1, b=2)
x=0 y=42 kwargs={'a':1, 'b':2}
\end{lstlisting}

  \+
  \ldots and also with variadic arguments:
\begin{lstlisting}
>>> def kwfn2(x, *rest, **kwargs):
...   print ("x=%s rest=%s kwargs=%s"
...          % (x, rest, kwargs))
>>> kwfn2(0, 1, 2, 3, a=1, b=2)
x=0 rest=(1, 2, 3) kwargs={'a':1, 'b':2}
\end{lstlisting}
\end{frame}


\begin{frame}[fragile,fragile]
  \frametitle{Keyword arguments, III}
  You can pass to a function a set of keyword arguments that is
  determined only at run time.

  The \texttt{**} operator takes any \emph{dictionary} and turns it
  into the bundle of keyword arguments for a function:
\begin{lstlisting}
>>> D = { 'c':4, 'd':2 }
>>> kwfn2(x=1, **D)
x=1 rest=() kwargs={ 'c':4, 'd':2 }
\end{lstlisting}
\end{frame}


\begin{frame}[fragile]
  \begin{exercise}
    Write a \texttt{maxarg} function, that takes arbitrary keyword
    arguments (but assume the values are all numbers), and returns the
    name of the argument with the largest value.

    \+
    Example:
\begin{lstlisting}
>>> maxarg(a=1, b=2)
'b'
\end{lstlisting}
  \end{exercise}
\end{frame}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
