\input{../shared/shared}

\renewcommand{\course}{Robotics}
\renewcommand{\coursepicture}{roboticsLecture}
\renewcommand{\coursedate}{Winter 2014}
\renewcommand{\topic}{SKIPPED THIS TERM -- Legged Locomotion (brief intro)}

\slides

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\slide{Legged Locomotion}{

\item Why legs?

~

Bacterial Flagellum: ~ (rotational ``motors'' in Biology?)

~

\centerline{
\showh[.4]{flagellum1}
\qquad
\showh[.4]{flagellum2}
}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\slide{Legged Locomotion}{

\item Why legs?

~

-- Human/Animal Locomotion Research
 
~

-- Potentially less weight

~

-- Better handling of rough terrains

~~ (climbing, isolated footholds, ladders, stairs)

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\slide{Rolling vs walking}{

~

\show[.45]{rollingVsWalking}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\slide{One-legged locomotion}{

\twocol{.4}{.5}{\center

\mov{\showh[.7]{3D_hopper}}{10-RoboticsLecture/3D_hopper_hopping.mpeg}

}{

\item Three separate controllers for:

-- hopping height

-- horizontal velocity ~ (foot placement)

-- attitude ~ (hip torques during stance)

~

\item Each a simple (PD-like) controller

}

~

~

{\tiny
Raibert et al.: \emph{Dynamically Stable Legged
  Locomotion.} 1985
\url{http://dspace.mit.edu/handle/1721.1/6820}

~

Tedrake: \emph{Applied Optimal Control for
Dynamically Stable Legged Locomotion.} PhD thesis (2004).
\url{http://groups.csail.mit.edu/robotics-center/public_papers/Tedrake04b.pdf}

}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\slide{Biped locomotion}{

\item Walking vs Running

Walking := in all instances at least one foot is on ground

Running := otherwise

~

\item 2 phases of Walking

-- double-support phase (in Robotics often statically stable)

-- single-support phase (statically instable)

\show[.4]{walkingFSC}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\slide{Asimo}{
\center

\hspace*{-8mm}
\twocol{.5}{.5}{
\mov{\show[.9]{11-Honda-All-New-ASIMO-hopping2}}{10-RoboticsLecture/11-Honda-All-New-ASIMO-hopping.mp4}
}{
\mov{\show[.9]{11-Honda-All-New-ASIMO-slowMotion}}{10-RoboticsLecture/11-Honda-All-New-ASIMO-slowMotion.mp4}
}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\key{Statically stable walk}
\slide{Statically Stable Walk}{

\item You could rest (hold pose) at any point in time and not fall over

~

\centerline{$\iff$ ~ \emph{CoG projected on ground is within \color{red}{support polygon}}}

~

CoG = center of gravity of all body masses

support polygon = hull of foot contact points

~

\item Try yourself: Move as slow as you can but make normal length
  steps...

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\key{Zero moment point (ZMP)}
\slide{Zero Moment Point}{

\item Vukobratovic's view on walking, leading to ZMP ideas:

``Basic characteristics of all biped locomotion systems are:

~

(i) the possibility of rotation of the overall system about one of the foot
edges caused by strong disturbances, which is equivalent to the appearance of an
unpowered (passive) DOF,

~

(ii) gait repeatability (symmetry), which is related to
regular gait only

~

(iii) regular interchangeability of single- and double-support
phases''

~

~

\tiny

Vukobratovic \& Borovac: \emph{Zero-moment point---Thirty five years
  of its life.} International Journal of Humanoid Robotics 1, 157-173,
  2004.
\url{http://www.cs.cmu.edu/~cga/legs/vukobratovic.pdf}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\slide{Zero Moment Point}{

\show[.5]{zmp1}

\show[.7]{zmp3}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\slide{Zero Moment Point}{

\item ``ZMP is defined as that point on the ground at
which the net moment of the inertial forces and the gravity forces has no
component along the horizontal axes.''
(Vukobratovic \& Borovac, 2004)

~

\item $f_i$ = force vector acting on body $i$ (gravity plus external)

$w_i$ = angular vel vector of body $i$

$I_i$ = interia tensor ($\in \RRR^{3\times 3}$) of body  $i$

$r_i = p_i - p_{\text{ZMP}}$ = relative position of body $i$ to ZMP

~

\item Definition: $p_{\text{ZMP}}$ is the point on the ground for which

\centerline{\Large
$
\sum_i r_i \times f_i + I_i \dot w_i + w_i \times I_i w_i
\overset{!}= (0,0,\*)^\T
$}

~

\tiny See also: Popovic, Goswami \& Herr: \emph{Ground Reference Points in Legged
  Locomotion: Definitions, Biological Trajectories and Control
  Implications.} International Journal of Robotics Research 24(12),
2005.
\url{http://www.cs.cmu.edu/~cga/legs/Popovic_Goswami_Herr_IJRR_Dec_2005.pdf}


}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\slide{Zero Moment Point}{

\item If the ZMP is outside the support polygon $\to$ foot rolls over
  the edge

  (presumes foot is a rigid body, and non-compliant control)

~

\item Locomotion is \emph{dynamically stable} if the ZMP remains within the
  foot-print polygons.

($\oto$ the support can always apply some momentum, if necessary)

~

\item Computing the ZMP in practise:
\anchor{10,-70}{\showh[.5]{zmp2}}

-- either exact robot model

-- or foot pressure sensors

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\slide{Zero Moment Point -- example}{

\item combine ZMP with 3D linear inverted pendulum model and
  model-predictive control

~

\showh[.4]{hrp-2}
\qquad
\showh[.25]{hrp-2-ZMPwalk}

\mov{HRP-2 stair climbing}{10-RoboticsLecture/HRP-2_Stair_Climbing.mp4}

~

\tiny

Kajita et al.: \emph{Biped Walking Pattern Generation by using Preview
Control of Zero-Moment Point.} ICRA 2003.
\url{http://eref.uqu.edu.sa/files/eref2/folder1/biped_walking_pattern_generation_by_usin_53925.pdf}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \slide{Zero Moment Point}{

%% Another commonly used approach:

%% \item Start with a hand-designed ro human motion capture trajectory.

%% \item Modify/optimize trajectory such that locomotion is stable
%%   according to the ZMP criterion.

%% \item Add online stabilization to deal with perturbations.

%% }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\slide{ZMP Summary}{

\item ZMP is the ``rescue'' for conventional methods:

-- ZMP-stability $\to$ the robot can be controlled as if foot is
   ``glued'' (virtually) to the ground!

-- The whole body behaves like a ``conventional arm''

-- Can accellerate $\ddot q$ any DoF $\to$ conventional dynamic control

\centerline{\emph{fully actuated system}}

~

\item Limitations:

-- Humans don't use ZMP stability, we allow our feet to roll

~~ (toe-off, heel-strike: ZMP at edge of support polygon)

-- Can't describe robots with point feet (walking on stilts)

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\key{Human locomotion}
\slide{Models of human bipedal locomotion}{

~

The following illustrations are from:

~

McMahon: \emph{Mechanics of Locomotion.} IJRR 3:4-28, 1984

{\tiny

\url{http://www.cs.cmu.edu/~cga/legs/mcmahon1.pdf}

\url{http://www.cs.cmu.edu/~cga/legs/mcmahon2.pdf}

}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\slide{}{

Walking research from Marey 1874:

~

\show[.4]{humanWalking1}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\slide{Six determinants of gait}{

{\small following Saunders, Inman \& Eberhart (1953)}

\only<1>{1. Compass Gait:\\[2ex] \show[.5]{humanWalking2}}
\only<2>{2. Pelvic Rotation:\\[2ex] \show[.55]{humanWalking2b}}
\only<3>{3. Pelvic Tilt:\\[2ex] \show[.5]{humanWalking3} \show[.5]{humanWalking3b}}
\only<4>{4. Stance Knee Flexion:\\[2ex] \show[.5]{humanWalking4}}
\only<5>{5. Stance Ankle Flextion:\\[2ex] \show[.5]{humanWalking5}}
\only<6>{6. Pelvis Lateral Displacement:\\[2ex] \show[.3]{humanWalking6}}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\slide{Models of human bipedal locomotion}{

\item Human model with 23 DoFs, 54 muscles

-- Compare human walking data with model

-- Model: optimize energy-per-distance 

-- Energy estimated based on metabolism and muscle heat rate models

~

\twocol{.5}{.5}{
\show[.25]{humanWalkModels}
}{
\mov{\show[.7]{walkingAnderson}}{10-RoboticsLecture/walkingAnderson.mpg}
}

~

~

\tiny

Anderson \& Pandy: \emph{Dynamic Optimization of Human
Walking.} Journal of Biomechanical Engineering 123:381-390, 2001.
\url{http://e.guigon.free.fr/rsc/article/AndersonPandy01.pdf}

Anderson \& Pandy:
\emph{Static and dynamic optimization solutions for gait
are practically equivalent.} Journal of Biomechanics 34 (2001) 153-161.
\url{http://www.bme.utexas.edu/faculty/pandy/StaticOptWalking2001.pdf}

}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\slide{Models of human bipedal locomotion}{

\item Suggest different principles of human motion:

-- passive dynamics (Compass Gait) ~ $\oto$ ~ underactuated system

-- modulation of basic passive dynamics

-- Energy minimization

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\key{Passive dynamic walking: Compass Gait}
\slide{Passive dynamic walking: Compass Gait}{

\item Basic 2D planar model of the Compass Gait:

\show[.4]{compassGait1}

~

The pose is described by $q=(\t_{s}, \t_{ns})$, the state by $(q, \dot q)$

~

\tiny
Goswami, Thuilot \& Espiau: \emph{A study of the passive gait of a
  compass-like biped robot: symmetry and chaos.} International Journal
of Robotics Research 17, 1998.

\url{http://www.ambarish.com/paper/COMPASS_IJRR_Goswami.pdf}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\slide{Passive dynamic walking: Compass Gait}{

\item Swing phase has analytic equations of motions

\centerline{\large $M(q) \ddot q + C(q,\dot q) \dot q + G(q) = 0$}

but can't be solved analytically...

\item Phase space plot of numeric solution:

\show[.6]{compassGait2}

}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\slide{Passive walker examples}{

\mov{compass gait simulation}{10-RoboticsLecture/compassGait1.mp4}

\mov{controlled on a circle}{10-RoboticsLecture/compassGait2-manipulated.flv}

\mov{passive walker}{10-RoboticsLecture/Passive_Dynamic_Walker.mp4}

~

\item Minimally actuated: 
\mov{Minimal Control on rough terrain}{10-RoboticsLecture/tedrakeCompassGait.avi}



}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\key{Impact models}
\slide{Impact Models in the Compass Gait}{

\item Switch between two consecutive swing phases: depends on slope!

\item Typical assumptions made in simulation models:

{\small

-- The contact of the swing leg with the ground results in
no rebound and no slipping of the swing leg.

-- At the moment of impact, the stance leg lifts from the
ground without interaction.

-- The impact is instantaneous.

-- The external forces during the impact can be represented
by impulses.

-- The impulsive forces may result in an instantaneous
change in the velocities, but there is no instantaneous
change in the configuration.

-- The actuators cannot generate impulses and, hence, can
be ignored during impact.
}

~

\tiny

Westervelt, Grizzle \& Koditschek: \emph{Hybrid Zero Dynamics of
  Planar Biped Walkers.}  IEEE Trans.\ on Automatic Control 48(1),
2003.

\url{http://repository.upenn.edu/cgi/viewcontent.cgi?article=1124&context=ese_papers}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\slide{Implausibility of the stiff Compass Gait leg}{

\show[.5]{walkingVsRunning}

~

~

\tiny

Geyer, Seyfarth \& Blickhan: \emph{Compliant leg behaviour explains
basic dynamics of walking and running.} Proc. Roy. Soc. Lond. B,
273(1603): 2861-2867, 2006.
\url{http://www.cs.cmu.edu/~cga/legs/GeyerEA06RoySocBiolSci.pdf}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\slide{Learning to walk in 20 Minutes}{

\item Policy Gradient method (Reinforcement Learning)

Stationary policy parameterized as linear in features $u = \sum_i w_i
\phi_i(q,\dot q)$

\item Problem: find parameters $w_i$ to minimize expected costs

cost = mimick $(q,\dot q)$ of the passive down-hill walker at
``certain point in cycle''

\centerline{\showh[.2]{tedrake-LearningToWalk}
\qquad
\mov{Learning To Walk}{10-RoboticsLecture/tedrake-LearningToWalk.avi}
}

~

\tiny

Tedrake, Zhang \& Seung: \emph{Stochastic
policy gradient reinforcement learning on a simple 3D biped.} IROS,
2849-2854, 2004.
\url{http://groups.csail.mit.edu/robotics-center/public_papers/Tedrake04a.pdf}

}

%% FROM RUSS:

%% ** WHat makes walking control hard?

%% -- Dynamic constraints (friction cones)

%%  - Underactuated (can't follow arbitrary trajectories)

%%  - Power limits at the actuators (Torque/Speed limits)

%%  - limits on ground reaction forces from the friction cones

%%  - fairly complicated constraints to reason about in the control system


%% ** What works:

%% -- trajectory optimization (can handle constraints)

%% -- Local stabalization of such trajectories

%%  - via Hybrid Zero Dynamics, or Transverse Linearization

%%  - more difficult to handle constraints (but only locally)

%%  - model-predictive control


%% -- Global path finding

%% -- Estimating regions of attraction



%% ** Rolling vs. Walking
%% \show{snakeRobot}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\slide{Summary}{\label{lastpage}

\item ZMP type walking was successful (ASIMO, HRP-2, etc), but limited

~

\item Future types of walking:

-- Exploit passive dynamics, cope with \emph{underactuation}

-- Follow some general optimiality principle ~ (but real-time!)

-- Learn ~ (esp.\ Reinforcement Learning)

-- Compliant hardware! ~ (controllable elasticity \& damping)

~

\item Recommended reading: Tedrake: \emph{Underactuated Robotics:
  Learning, Planning, and Control for Effcient and Agile Machines.}
  Course Notes for MIT 6.832

{\tiny\url{www.cs.berkeley.edu/~pabbeel/cs287-fa09/readings/Tedrake-Aug09.pdf}}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\slide{Finally, multi-legged locomotion}{

~

\show{gaits}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\slide{Finally, multi-legged locomotion}{

~

\mov{\show[.3]{2010-BostonDynamics-RISE}}{10-RoboticsLecture/2010-BostonDynamics-RISE.wmv}

~

\tiny
\url{http://www.bostondynamics.com/robot_rise.html}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\slidesfoot


%% COURSE PAGES
%% Atkeson:
%% http://www.cc.gatech.edu/fac/Chris.Atkeson/legs/
%% http://www.cs.cmu.edu/~cga/legs/
%% EPFL:
%% http://moodle.epfl.ch/course/view.php?id=261
%% MIT:
%% http://www.ai.mit.edu/projects/leglab/6894/6894.html


%% ****
%% NOTES from Christ Atkeson's lectures: topic ideas!

%% http://www.cc.gatech.edu/fac/Chris.Atkeson/legs/topics.txt

%% Topics:
%% *** indicates "classic, must read"
%% +++ indicates "new interesting stuff"

%% --------------------------------------------------------------------
%% Understanding Task, Mechanics

%% - biomechanics
%%   *** McMahon paper or book chapter
%%   *** Pandy and Anderson

%% - passive dynamic walking
%%   *** McGeer
%%   simplest (compass w/ point feet)
%%   extended feet
%%   knees
%%   body
%%   3D
%%   role of double support

%% - powered passive dynamic walking

%% - actuators and compliance
%%   - muscle
%%   - M2

%% - non anthropomorphic designs

%% --------------------------------------------------------------------
%% Existing systems

%% - bipeds
%%   ZMP-based: Honda, Sony, HRP-2, H7, ...
%%   Raibert
%%   Pratt - virtual control

%% - quadrupeds
%%   AIBO

%% - hexapods
%%   Rhex

%% --------------------------------------------------------------------
%% Control Design

%% - manual approaches
%%   Hodgins walk-run paper

%% - control using dynamic systems
%%   Kimura

%% - nonlinear control design
%%   Hopping robot controller papers series
%%   Grizzle, UMich

%% - "understanding dynamics
%%   abstraction/simplification

%% - trajectory optimization
%%   understanding human gait as well as generating robot gait
%%   *** Pandy and Anderson

%% - policy optimization
%%   simulated annealing
%%   evolutionary approaches (graphics, robotics, ...)
%%   policy search in machine learning, including derivative based.

%% --------------------------------------------------------------------
%% Robust Control and Disturbances

%% - How design a robust controller (wrt initial states, disturbances,
%% and modeling error)?

%% - How handle wide range of activities (sit, stand, avoid obstacles,
%% stepping stones, carry loads, ...)

%% - How handle wide range of disturbances (slip, trip, stumble;
%% respond with continuous response, recovery step(s), using arms or
%% grabbing support, safe fall)
%%   Hodgins paper(s)
%%   Ukemi paper

%% --------------------------------------------------------------------
%% Measurement

%% - how deal with measurement problem
%%   state estimation
%%   reduced order controllers

%% --------------------------------------------------------------------
%% Planning for Dynamic Behaviors

%% - Extending ZMP approach
%%   filters

%% - Kuffner stuff

%% - data driven behavior
%%   style

%% - templates (Koditschek)

%% - rough terrain

%% - coordinating lower and upper body
%%   pick up, carry, move, put down
%%   grab for support
%%   wave arms to assist balance

%% - coordinating behaviors
%%   soccer (kick and run)

%% - gait transitions, supporting multiple gaits
%%   kinematics
%%   mechanism

%% - autonomy

%% --------------------------------------------------------------------_
%% Intermittent Control Not Involving Legs

%% - snake locomotion

%% - juggling/manipulation

%% --------------------------------------------------------------------
%% Other topics

%% - scaling

%% - fall prediction and balance assessment in humans

%% - learning?

%% - dancing

%% - graphical characters
%%   interactive
%%   non-interactive
%% --------------------------------------------------------------------
%% --------------------------------------------------------------------

