\input{../shared/shared}

\renewcommand{\course}{Robotics}
\renewcommand{\coursepicture}{roboticsLecture}
\renewcommand{\coursedate}{Winter 2014}
\renewcommand{\exnum}{5}

\exercises

In the lecture we discussed PD force control on a 1D point mass,
which leads to oscillatory behavior for high $K_p$ and damped behavior
for high $K_d$. Slide 03:15 replaces the parameters
$K_p,K_d$ by two other, more intuitive parameters, $\lambda$ and
$\xi$: $\lambda$ roughly denotes the time (or time steps) until the
goal is reached, and $\xi$ whether it is reached agressively ($\xi>1$,
which overshoots a bit) or by exponential decay ($\xi\le 1$). Use this
to solve the following exercise.

\exsection{PD force control on a 1D mass point}

a) Implement the system equation for a 1D point mass with mass
$m=3.456$. That is, implement the Euler intergration (see below) of
the system dynamics that computes $x_{t\po}$ given $x_t$ and $u_t$ in
each iteration, where $x_t = (q_t,\dot q_t)$.

There is no need for the robot simulator code---implement it directly
in plain C++ or another language.

Assume a step time of $\tau=0.01$sec. Generate a trajectory from the
start position $q_0=0$ and velocity $\dot q_0=0$ that approaches the
goal position $q^*=1$ with high precision within about 1 second using
PD force control.  Find 3 different parameter sets for $K_p$ and $K_d$
to get oscillatory, overdamped and critical damped behaviors. Plot the
point trajectory (e.g.\ using the routine
%
@gnuplot(arr& q); MT::wait();@ .)

b) Repeat for time horizon $t=2$sec and $t=5$sec. How should the
values of $K_p$ and $K_d$ change when we have more time?

c) Implement a PID controller (including the integral (stationary error)
term). How does the solution behave with only $K_i$ turned on
($K_p=K_d=0$); how with $K_i$ and $K_d$ non-zero?

\textbf{Note on Euler Integration:} See
also \url{http://en.wikipedia.org/wiki/Euler_method}. Given a
differential equation $\dot x = f(x,u)$, Euler integration numerically
``simulates'' this equation forward by iterating
$$x_{t\po} = x_t + \tau f(x,u)$$
for small time steps $\tau$.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\exsection{Following a reference trajectory with PD control}

We have the same point mass as above. But now the goal position $q^*$
changes over time: we have a reference trajectory
$$q^*(t) = \cos(t)$$

How can one use a PD controller to let the point mass
robustly and \emph{precisely} follow this reference trajectory?

Simulate the problem as above, but at every full second apply a randomized
impuls (=instantaneous change in velocity) on the point mass. That
is, simply set $\dot q \gets \dot q + \xi$ for $\xi\sim\NN(0,.1)$. (In
code: @double xi = 0.1 * rnd.gauss();@) Generate oscillatory,
overdamped and critical recovery behaviors.


%% A PD controller can not only be used to stabalize a system around a
%% fixed target $q^*,\dot q^*$. It can also be used to follow a moving
%% target $q^*_t$, which is called a \emph{reference trajectory}. From a
%% given a reference trajectory $q^\text{ref}_{0:T}$ of joint angles only, we
%% compute the velocity as $\dot q^\text{ref}_t=(q_{t\po}-q_{t-1})/(2\tau)$ and the
%% acceleration as $\ddot q_t^\text{ref} = (q_{t\1} + q_{t\po} -
%% 2q_t)/\tau^2$. The last equation on slide 04:22 shows how we can use
%% PD acceleration control to follow such a reference trajectory. Use
%% this to finally solve the full-fleged, noisy, dynamic peg-in-a-hole
%% problem in the next exercise.



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\exerfoot
